---
title: "GA4 data preparation"
subtitle: "Import, standardisation and reusable clean outputs"
author: "Jordan"
date: today
format:
  html:
    toc: true
    number-sections: true
    code-fold: true
    embed-resources: true
execute:
  warning: false
  message: false
---

# Purpose

This report is the single preparation stage for the GA4 pipeline. It imports
the five supplied GA4 exports, validates the expected variables, cleans names
and numeric fields, standardises page paths, adds the common reporting period,
and writes analysis-ready CSV files. Raw exports are never overwritten.

# Setup and source contract

```{r}
library(tidyverse)
library(janitor)
library(here)

options(scipen = 999)
here::i_am("reports/GA4/01_ga4_data_preparation.qmd")

raw_dir <- here("data", "raw", "GA4")
clean_dir <- here("data", "clean", "GA4")
analysis_start_date <- as.Date("2025-07-01")
analysis_end_date <- as.Date("2026-06-30")

source_files <- c(
  demographic = "Demographic_details_Country.csv",
  events = "Events_Event_name.csv",
  landing_pages = "Landing_page_Landing_page.csv",
  pages = "Pages_and_screens_Page_path_and_screen_class.csv",
  traffic = "Traffic_acquisition_Session_primary_channel_group_(Default_Channel_Group).csv"
)

stopifnot(all(file.exists(file.path(raw_dir, source_files))))
```

# Import and select variables

GA4 CSVs contain nine report-metadata rows before the header. All analytical
columns are imported as text first so type conversion is explicit and auditable.

```{r}
read_ga4 <- function(file) {
  read_csv(
    file.path(raw_dir, file), skip = 9,
    col_types = cols(.default = col_character()),
    trim_ws = TRUE, show_col_types = FALSE
  ) |>
    clean_names() |>
    mutate(across(where(is.character), ~ na_if(str_squish(.x), "")))
}

ga4_raw <- map(source_files, read_ga4)

required_fields <- list(
  demographic = c("country", "active_users", "new_users", "engaged_sessions",
    "engagement_rate", "engaged_sessions_per_active_user",
    "average_engagement_time_per_active_user", "event_count", "key_events",
    "user_key_event_rate", "total_revenue"),
  events = c("event_name", "event_count", "total_users",
    "event_count_per_active_user", "total_revenue"),
  landing_pages = c("landing_page", "sessions", "active_users", "new_users",
    "average_engagement_time_per_session", "key_events", "total_revenue",
    "session_key_event_rate"),
  pages = c("page_path_and_screen_class", "views", "active_users",
    "views_per_active_user", "average_engagement_time_per_active_user",
    "event_count", "key_events", "total_revenue"),
  traffic = c("session_primary_channel_group_default_channel_group", "sessions",
    "engaged_sessions", "engagement_rate", "average_engagement_time_per_session",
    "events_per_session", "event_count", "key_events", "session_key_event_rate",
    "total_revenue")
)

walk2(ga4_raw, required_fields, ~ stopifnot(all(.y %in% names(.x))))
ga4_selected <- map2(ga4_raw, required_fields, ~ select(.x, all_of(.y)))
```

# Clean dates and numeric fields

The dates are report-level metadata, not row-level observations. Consequently,
these files support full-period summaries but not daily or monthly trends.

```{r}
parse_number_strict <- function(x) readr::parse_number(
  x, na = c("", "NA"), locale = locale(grouping_mark = ",")
)

dimension_fields <- c(
  "country", "event_name", "landing_page",
  "page_path_and_screen_class",
  "session_primary_channel_group_default_channel_group"
)

ga4_typed <- map(
  ga4_selected,
  ~ mutate(.x, across(-any_of(dimension_fields), parse_number_strict))
)

add_period <- function(x) mutate(
  x,
  analysis_start_date = analysis_start_date,
  analysis_end_date = analysis_end_date
)
```

# Create analysis-ready tables

```{r}
standardise_path <- function(x) {
  out <- x |> str_squish() |> str_remove("\\?.*$")
  if_else(out %in% c("/", "(not set)"), out, str_remove(out, "/+$"))
}

demographic_clean <- ga4_typed$demographic |>
  mutate(
    new_user_ratio = if_else(active_users > 0, new_users / active_users, NA_real_),
    review_flag = active_users >= 500 &
      (engagement_rate < 0.15 | average_engagement_time_per_active_user < 10)
  ) |>
  add_period()

events_clean <- ga4_typed$events |>
  mutate(
    event_category = case_when(
      event_name %in% c("page_view", "session_start", "first_visit",
        "user_engagement", "scroll") ~ "Standard/enhanced measurement",
      str_detect(event_name, regex("form|register|sign_up|purchase|member",
        ignore_case = TRUE)) ~ "Potential business conversion",
      str_detect(event_name, regex("click|download|file|video",
        ignore_case = TRUE)) ~ "Content interaction",
      TRUE ~ "Other"
    )
  ) |>
  add_period()

traffic_clean <- ga4_typed$traffic |>
  rename(channel = session_primary_channel_group_default_channel_group) |>
  mutate(
    session_share = sessions / sum(sessions, na.rm = TRUE),
    engagement_rate_check = engaged_sessions / sessions,
    engagement_rate_difference = abs(engagement_rate - engagement_rate_check)
  ) |>
  add_period()

landing_pages_clean <- ga4_typed$landing_pages |>
  rename(landing_page_original = landing_page) |>
  mutate(
    has_query_parameter = str_detect(landing_page_original, fixed("?")),
    landing_page_standardised = standardise_path(landing_page_original)
  ) |>
  add_period()

pages_clean <- ga4_typed$pages |>
  rename(page_path_original = page_path_and_screen_class) |>
  mutate(
    has_query_parameter = str_detect(page_path_original, fixed("?")),
    page_path_standardised = standardise_path(page_path_original),
    page_path_template = page_path_standardised |>
      str_replace_all("(/view-order/)\\d+", "\\1{id}") |>
      str_replace_all("(/view-subscription/)\\d+", "\\1{id}") |>
      str_replace_all("(/members-area/)\\d+", "\\1{id}") |>
      str_replace_all("(/groups/)\\d+", "\\1{id}") |>
      str_replace_all("(/join-group/)[A-Za-z0-9_-]{16,}", "\\1{token}"),
    page_path_was_templated = page_path_template != page_path_standardised
  ) |>
  add_period()
```

These exports have different grains (country, event, landing page, page path
and channel). They cannot be row-joined without inventing relationships. Their
valid merge is a manifest that links each table to the same reporting period.

```{r}
clean_outputs <- list(
  ga4_demographic_clean = demographic_clean,
  ga4_events_clean = events_clean,
  ga4_landing_pages_clean = landing_pages_clean,
  ga4_pages_and_screens_clean = pages_clean,
  ga4_traffic_acquisition_clean = traffic_clean
)

ga4_output_manifest <- imap_dfr(clean_outputs, ~ tibble(
  dataset = .y, file_name = paste0(.y, ".csv"),
  rows = nrow(.x), columns = ncol(.x),
  analysis_start_date = analysis_start_date,
  analysis_end_date = analysis_end_date
))

knitr::kable(ga4_output_manifest)
```

# Write clean data

```{r}
dir.create(clean_dir, recursive = TRUE, showWarnings = FALSE)
iwalk(clean_outputs, ~ write_csv(.x, file.path(clean_dir, paste0(.y, ".csv")), na = ""))
write_csv(ga4_output_manifest, file.path(clean_dir, "ga4_output_manifest.csv"), na = "")

ga4_cleaning_log <- tribble(
  ~step, ~rule, ~result,
  "Import", "Skip nine GA4 metadata rows", "Completed",
  "Variables", "Validate and retain analysis fields", "Completed",
  "Types", "Parse numeric fields explicitly", "Completed",
  "Dates", "Attach report-level start and end dates", "Completed",
  "Paths", "Preserve original paths and add standardised/template fields", "Completed",
  "Merge", "Link different grains through a common-period manifest", "Completed"
)
write_csv(ga4_cleaning_log, file.path(clean_dir, "ga4_cleaning_log.csv"))
knitr::kable(ga4_cleaning_log)
```
