---
title: "GA4 data quality"
subtitle: "Coverage, completeness, duplicates, continuity and metric checks"
author: "Jordan"
date: today
format:
  html: {toc: true, number-sections: true, code-fold: true, embed-resources: true}
execute: {warning: false, message: false}
---

# Purpose

This report checks the outputs created by `01_ga4_data_preparation.qmd`.
Potential issues are reported, not silently deleted or corrected.

# Setup

```{r}
library(tidyverse)
library(here)
library(scales)
here::i_am("reports/GA4/02_ga4_data_quality.qmd")

clean_dir <- here("data", "clean", "GA4")
files <- c(
  demographic = "ga4_demographic_clean.csv",
  events = "ga4_events_clean.csv",
  landing_pages = "ga4_landing_pages_clean.csv",
  pages = "ga4_pages_and_screens_clean.csv",
  traffic = "ga4_traffic_acquisition_clean.csv"
)
stopifnot(all(file.exists(file.path(clean_dir, files))))
ga4 <- map(files, ~ read_csv(file.path(clean_dir, .x), show_col_types = FALSE))
```

# Data range and structure

```{r}
structure_summary <- imap_dfr(ga4, ~ summarise(.x,
  dataset = .y, rows = n(), columns = ncol(.x),
  start_date = min(analysis_start_date), end_date = max(analysis_end_date)
))
knitr::kable(structure_summary)
```

# Missing values and duplicate records

```{r}
missing_summary <- imap_dfr(ga4, ~ tibble(
  dataset = .y,
  missing_cells = sum(is.na(.x)),
  fields_with_missing = sum(map_lgl(.x, ~ any(is.na(.x))))
))

key_fields <- c(
  demographic = "country", events = "event_name",
  landing_pages = "landing_page_original", pages = "page_path_original",
  traffic = "channel"
)

duplicate_summary <- imap_dfr(ga4, ~ tibble(
  dataset = .y,
  exact_duplicate_rows = sum(duplicated(.x)),
  duplicate_dimension_values = sum(duplicated(.x[[key_fields[[.y]]]]))
))

knitr::kable(missing_summary)
knitr::kable(duplicate_summary)
```

# Date continuity

Each row only carries the same report start and end dates. There is no daily or
monthly dimension, so continuity cannot be tested between those dates.

```{r}
date_quality <- tibble(
  check = c("Common start date", "Common end date", "Row-level date available",
    "Daily/monthly continuity test"),
  result = c(
    as.character(n_distinct(map_dfr(ga4, ~ select(.x, analysis_start_date))$analysis_start_date) == 1),
    as.character(n_distinct(map_dfr(ga4, ~ select(.x, analysis_end_date))$analysis_end_date) == 1),
    "No", "Not computable from current exports"
  )
)
knitr::kable(date_quality)
```

# Metric anomalies and reconciliation

```{r}
page_view_events <- ga4$events |>
  filter(event_name == "page_view") |>
  summarise(n = sum(event_count, na.rm = TRUE)) |>
  pull(n)

metric_checks <- tibble(
  check = c(
    "Negative numeric cells", "Engagement rates outside 0–1",
    "Engaged sessions above sessions", "Traffic engagement-rate differences",
    "Pages with views below active users", "All key-event totals are zero",
    "Pages views minus page_view events", "Flagged high-volume countries"
  ),
  result = c(
    sum(map_dbl(ga4, ~ sum(select(.x, where(is.numeric)) < 0, na.rm = TRUE))),
    sum(ga4$traffic$engagement_rate < 0 | ga4$traffic$engagement_rate > 1, na.rm = TRUE) +
      sum(ga4$demographic$engagement_rate < 0 | ga4$demographic$engagement_rate > 1, na.rm = TRUE),
    sum(ga4$traffic$engaged_sessions > ga4$traffic$sessions, na.rm = TRUE),
    sum(ga4$traffic$engagement_rate_difference > 1e-6, na.rm = TRUE),
    sum(ga4$pages$views < ga4$pages$active_users, na.rm = TRUE),
    as.numeric(sum(map_dbl(ga4, ~ {
      if ("key_events" %in% names(.x)) sum(.x$key_events, na.rm = TRUE) else 0
    })) == 0),
    sum(ga4$pages$views, na.rm = TRUE) - page_view_events,
    sum(ga4$demographic$review_flag, na.rm = TRUE)
  )
)
knitr::kable(metric_checks)
```

# Quality conclusion

The exports are suitable for aggregate traffic and content analysis. The main
limitations are analytical coverage rather than CSV corruption: no time series,
no source/medium detail, all Key Event values are zero, and some high-volume
country rows require investigation before geographic comparisons are trusted.
