---
title: "GA4 traffic analysis"
subtitle: "Users, sessions and acquisition"
author: "Jordan"
date: today
format:
  html: {toc: true, number-sections: true, code-fold: true, embed-resources: true}
execute: {warning: false, message: false}
---

# Setup

```{r}
library(tidyverse)
library(here)
library(scales)
here::i_am("reports/GA4/03_ga4_traffic_analysis.qmd")
read_clean <- function(x) read_csv(here("data", "clean", "GA4", x), show_col_types = FALSE)
demographic <- read_clean("ga4_demographic_clean.csv")
events <- read_clean("ga4_events_clean.csv")
traffic <- read_clean("ga4_traffic_acquisition_clean.csv")
```

# Users, sessions and views

User metrics are not additive across countries. The Events export provides the
most appropriate supplied overall user count; sessions come from acquisition
and views reconcile between Events and Pages and Screens.

```{r}
traffic_kpis <- tibble(
  metric = c("Users reporting page_view", "Sessions", "Views"),
  value = c(
    events |> filter(event_name == "page_view") |> pull(total_users),
    sum(traffic$sessions, na.rm = TRUE),
    events |> filter(event_name == "page_view") |> pull(event_count)
  )
)
knitr::kable(traffic_kpis)
```

# Traffic trend

A trend chart cannot be produced because all files contain one aggregate row
set for the full 12-month period. Future exports should add `date` or `month`
alongside users, sessions and views.

```{r}
trend_availability <- tibble(
  requested_output = "Monthly users, sessions and views trend",
  status = "Unavailable",
  reason = "No row-level date/month dimension in supplied exports",
  required_fields = "date or month; active users; sessions; views"
)
knitr::kable(trend_availability)
```

# Channel performance

```{r}
channel_table <- traffic |>
  transmute(
    channel, sessions, session_share = percent(session_share, accuracy = 0.1),
    engagement_rate = percent(engagement_rate, accuracy = 0.1),
    average_engagement_seconds = round(average_engagement_time_per_session, 1),
    events_per_session = round(events_per_session, 2), key_events
  ) |>
  arrange(desc(sessions))
knitr::kable(channel_table)

ggplot(traffic, aes(x = reorder(channel, sessions), y = sessions)) +
  geom_col(fill = "#2C7FB8") + coord_flip() +
  scale_y_continuous(labels = comma) +
  labs(x = NULL, y = "Sessions", title = "Sessions by default channel group") +
  theme_minimal()
```

# Source and medium

Only the default channel group was supplied. Source, medium and campaign
performance cannot be separated, and Direct or Unassigned traffic cannot be
diagnosed from this export. Add `session source`, `session medium` and
`session campaign` to the next acquisition export.

# New and returning users

```{r}
new_user_context <- demographic |>
  summarise(
    reported_country_rows = n(),
    sum_active_users_across_countries = sum(active_users),
    sum_new_users_across_countries = sum(new_users),
    countries_flagged_for_review = sum(review_flag)
  )
knitr::kable(new_user_context)
```

The country export includes New users but not Returning users. Subtracting New
users from Active users is not a valid replacement because GA4 user metrics
have different definitions and country rows can double-count people. A future
export should use the `new / established` dimension (or an agreed equivalent)
with user counts and date/month.
