---
title: "GA4 summary and recommendations"
subtitle: "Integrated findings, implications and next actions"
author: "Jordan"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    embed-resources: true
execute: {warning: false, message: false, echo: false}
---

```{r, include=FALSE}
library(tidyverse)
library(here)
library(scales)
options(scipen = 999)
here::i_am("reports/GA4/05_ga4_summary_recommendations.qmd")

read_clean <- function(x) read_csv(here("data", "clean", "GA4", x), show_col_types = FALSE)
demographic <- read_clean("ga4_demographic_clean.csv")
events <- read_clean("ga4_events_clean.csv")
landing <- read_clean("ga4_landing_pages_clean.csv")
pages <- read_clean("ga4_pages_and_screens_clean.csv")
traffic <- read_clean("ga4_traffic_acquisition_clean.csv")

total_sessions <- sum(traffic$sessions)
total_engaged <- sum(traffic$engaged_sessions)
overall_er <- total_engaged / total_sessions
weighted_time <- weighted.mean(traffic$average_engagement_time_per_session, traffic$sessions)
total_views <- sum(pages$views)
total_events <- sum(traffic$event_count)
page_view <- events |> filter(event_name == "page_view")
direct <- traffic |> filter(channel == "Direct")
organic <- traffic |> filter(channel == "Organic Search")
referral <- traffic |> filter(channel == "Referral")
ai_assistant <- traffic |> filter(channel == "AI Assistant")
email_sessions <- sum(traffic$sessions[traffic$channel == "Email"])
landing_total <- sum(landing$sessions)
not_set_sessions <- sum(landing$sessions[landing$landing_page_original == "(not set)"])
not_set_share <- not_set_sessions / landing_total
session_difference <- total_sessions - landing_total
# Select one report for the headline total. Adding the same metric across
# reports would double-count it because each report is a different breakdown.
reported_key_events <- sum(traffic$key_events)
key_event_reconciliation <- tibble(
  report = c("Traffic acquisition", "Landing pages", "Pages and screens", "Country"),
  key_events = c(sum(traffic$key_events), sum(landing$key_events),
    sum(pages$key_events), sum(demographic$key_events))
)
scroll_events <- events |> filter(event_name == "scroll")
scroll_to_view_ratio <- scroll_events$event_count / total_views
templated_rows <- sum(pages$page_path_was_templated)
review_pages <- pages |>
  filter(views >= 100, average_engagement_time_per_active_user < 10,
         page_path_standardised != "(not set)") |>
  arrange(average_engagement_time_per_active_user, desc(views))
flagged_countries <- demographic |> filter(review_flag) |> arrange(desc(active_users))
```

# Quick guide to the measures {.unnumbered}

Use this table as a glossary while reading the report. Counts at different
levels should not be added together.

```{r}
measure_guide <- tribble(
  ~measure, ~short_description,
  "Total users", "Distinct users GA4 identified as triggering the relevant event during the period",
  "Active users", "Users GA4 classified as engaged or active during the period",
  "New users", "Users GA4 identified as visiting for the first time",
  "Sessions", "Continuous visits; one user can generate more than one session",
  "Engaged sessions", "Sessions lasting over 10 seconds, containing a Key Event, or containing 2+ page views",
  "Engagement rate", "Engaged sessions divided by all sessions",
  "Average engagement seconds per session", "Average active engagement time recorded for one visit",
  "Landing page", "The first page of a session",
  "Page views / Views", "Times pages were viewed, including repeat views and refreshes",
  "Views per active user", "Page views divided by active users for that page",
  "Average engagement seconds per active user", "Page engagement time divided by active users for that page",
  "Event count", "Total recorded interactions such as page_view, scroll or click",
  "Events per session", "Event count divided by sessions",
  "Key Events", "Events configured to represent important business outcomes",
  "Channel", "GA4 category describing how a session arrived, such as Direct or Organic Search",
  "Share of all sessions", "A channel's sessions divided by sessions across all channels",
  "(not set)", "GA4 could not determine a value for the dimension"
)
knitr::kable(measure_guide, caption = "Short descriptions of measures used in this report")
```

# Executive summary

```{r}
headline_kpis <- tibble(
  measure = c("Total users who triggered page_view", "Sessions", "Page views",
    "Engaged sessions", "Overall engagement rate",
    "Average engagement time/session", "Events", "Reported Key Events"),
  result = c(comma(page_view$total_users), comma(total_sessions), comma(total_views),
    comma(total_engaged), percent(overall_er, 0.1),
    paste0(number(weighted_time, 0.1), " seconds"), comma(total_events),
    comma(reported_key_events)),
  short_description = c(
    "Users who triggered at least one page_view",
    "Continuous website visits",
    "Page loads and repeat views",
    "Sessions meeting at least one GA4 engagement condition",
    "Share of sessions classified as engaged",
    "Average active engagement time for each session",
    "All recorded website interactions",
    "Tracked business outcomes available in these exports")
)
knitr::kable(
  headline_kpis,
  caption = "Headline website activity",
  align = c("l", "r", "l")
)
```

## Reading the headline numbers

Read the four activity measures as a sequence, not as numbers to add together:

```text
55,603 users
→ made 92,227 visits (sessions)
→ viewed 193,799 pages
→ generated 605,209 recorded interactions (events)
```

One user can visit more than once, one visit can contain several page views,
and one page view can generate several events. For example, opening a page,
scrolling near the bottom and clicking an external link can create one page view
and several events within the same session.

The reporting period is **`r format(min(traffic$analysis_start_date), '%d %B %Y')`
to `r format(max(traffic$analysis_end_date), '%d %B %Y')`**. Of the
**`r comma(total_sessions)` sessions**, **`r comma(total_engaged)`** met at least
one GA4 engagement condition. This produces an overall engagement rate of
**`r percent(overall_er, 0.1)`**. Average recorded engagement time was
**`r number(weighted_time, 0.1)` seconds per session**.

The website attracts substantial interest in Science Talent Search, member
resources, events and publications. Organic Search is the strongest large
channel on engagement. Direct supplies the most sessions but engages materially
less strongly and may contain untagged campaign traffic. The central weakness
is measurement rather than lack of activity: form interactions exist, but no
supplied report contains a non-zero Key Event. STAV therefore cannot yet connect
traffic and content consumption to registrations, memberships or other outcomes.

# Where visits came from (Acquisition)

The table below provides the common evidence base for the channel findings that
follow. **Share of all sessions** is calculated as:

```text
Sessions for the channel ÷ Sessions across all channels
```

It measures traffic contribution, not user share, engagement quality or
conversion performance.

```{r}
traffic |>
  arrange(desc(sessions)) |>
  transmute(
    channel,
    sessions = comma(sessions),
    `share of all sessions` = percent(session_share, 0.1),
    engagement_rate = percent(engagement_rate, 0.1),
    `average engagement seconds per session` = number(
      average_engagement_time_per_session, 0.1
    ),
    events_per_session = number(events_per_session, 0.01)
  ) |>
  knitr::kable(
    caption = "Channel comparison used in Sections 2.1–2.3",
    align = c("l", "r", "r", "r", "r", "r")
  )
```

::: {.callout-note title="How engagement rate is calculated and recorded"}
**Engagement rate = engaged sessions / all sessions.** GA4 marks a session as
engaged when it lasts more than 10 seconds, contains at least one Key Event, or
contains at least two page or screen views. The website tag sends events and
engagement-time information to GA4; GA4 then determines whether each session
meets at least one of these conditions. It is a session-quality indicator, not
a conversion rate or a direct measure of satisfaction.
:::

## Direct brings scale, but attribution and engagement need attention

Direct generated **`r comma(direct$sessions)` sessions
(`r percent(direct$session_share, 0.1)`)**, making it the largest channel. Its
engagement rate was **`r percent(direct$engagement_rate, 0.1)`**, compared with
**`r percent(organic$engagement_rate, 0.1)`** for Organic Search. Average
engagement time was **`r number(direct$average_engagement_time_per_session, 0.1)`
seconds** for Direct versus **`r number(organic$average_engagement_time_per_session,
0.1)` seconds** for Organic Search.

This does not prove that Direct visitors are less valuable. With no source,
medium or campaign fields, Direct may include email, social, partner or offline
links missing campaign parameters. A **`r percent(direct$session_share, 0.1)`
Direct share** combined with only **`r comma(email_sessions)` Email sessions**
is a strong reason to audit campaign tagging.

## Organic Search is the strongest established acquisition channel

Organic Search contributed **`r comma(organic$sessions)` sessions
(`r percent(organic$session_share, 0.1)`)** with a
**`r percent(organic$engagement_rate, 0.1)` engagement rate**. Referral was
smaller at **`r comma(referral$sessions)` sessions**, but achieved a similar
**`r percent(referral$engagement_rate, 0.1)` engagement rate**. Search visibility
and relevant external links appear important and should be protected during
website or URL changes.

## AI Assistant is visible but currently very small

AI Assistant generated **`r comma(ai_assistant$sessions)` sessions**, equal to
**`r percent(ai_assistant$session_share, 0.01)` of all sessions**. GA4 uses this
channel when a recognised assistant such as ChatGPT, Gemini, DeepSeek, Copilot
or Grok refers the visit. At this volume, its engagement rate is descriptive but
not stable enough for strong performance conclusions.

Growth depends mainly on making STAV content easy for assistants to discover,
understand and cite: publish clear authoritative answers, use descriptive page
titles and headings, keep important information in crawlable HTML, maintain
accurate event and resource pages, add relevant structured data, and earn links
and mentions from trusted education organisations. Source/medium and landing-page
detail should be exported so future reports can identify which assistants and
which cited pages actually generate visits.

```{r}
traffic |>
  mutate(channel = fct_reorder(channel, sessions)) |>
  ggplot(aes(channel, sessions, fill = engagement_rate)) +
  geom_col() + coord_flip() + scale_y_continuous(labels = comma) +
  scale_fill_gradient(low = "#BFD7EA", high = "#005A8D", labels = percent) +
  labs(title = "Direct and Organic Search dominate reported sessions",
    subtitle = "Fill colour represents engagement rate", x = NULL,
    y = "Sessions", fill = "Engagement rate") + theme_minimal()
```

# Which pages people entered and viewed

## Science Talent Search and the homepage are the main entry points

```{r}
landing |>
  arrange(desc(sessions)) |>
  transmute(
    landing_page = landing_page_standardised,
    sessions = comma(sessions),
    active_users = comma(active_users),
    new_users = comma(new_users),
    `average engagement seconds per session` = number(
      average_engagement_time_per_session, 0.1
    )
  ) |>
  slice_head(n = 12) |>
  knitr::kable(
    caption = "Top landing pages: the first page of each session",
    align = c("l", "r", "r", "r", "r")
  )
```

The first table ranks the pages where sessions began. The next table answers a
different question: which pages received the most views at any point in a
session. A page can be highly viewed without being a leading landing page.

```{r}
pages |>
  filter(page_path_standardised != "(not set)") |>
  arrange(desc(views)) |>
  transmute(page = page_path_standardised, views = comma(views),
    active_users = comma(active_users),
    views_per_active_user = number(views_per_active_user, 0.01),
    `average engagement seconds per active user` = number(
      average_engagement_time_per_active_user, 0.1
    )) |>
  slice_head(n = 12) |>
  knitr::kable(
    caption = "Top content pages: views occurring anywhere in a session",
    align = c("l", "r", "r", "r", "r")
  )
```

`/science-talent-search` was the leading landing page with
**`r comma(landing$sessions[landing$landing_page_standardised == '/science-talent-search'][1])`
sessions**. The homepage followed with
**`r comma(landing$sessions[landing$landing_page_standardised == '/'][1])`
sessions**. Leading content also includes `/my-account`, `/sts-resources`, event
pages and trial-exam content, showing demand across public programmes,
professional resources and member-related activity.

High traffic shows content demand, not business success. Because Key Events are
zero, these tables cannot show whether visits resulted in registrations,
purchases or membership actions.

## Why unidentified landing pages matter

```{r}
landing_status <- landing |>
  mutate(status = if_else(landing_page_original == "(not set)",
    "Landing page not identified", "Landing page identified")) |>
  group_by(status) |>
  summarise(
    average_engagement_seconds_per_session = weighted.mean(
      average_engagement_time_per_session, sessions
    ),
    sessions = sum(sessions),
    active_users_across_rows = sum(active_users),
    new_users_across_rows = sum(new_users),
    .groups = "drop"
  ) |>
  mutate(
    share_of_landing_page_sessions = sessions / sum(sessions),
    sessions = comma(sessions),
    share_of_landing_page_sessions = percent(
      share_of_landing_page_sessions, 0.1
    ),
    active_users_across_rows = comma(active_users_across_rows),
    new_users_across_rows = comma(new_users_across_rows),
    average_engagement_seconds_per_session = number(
      average_engagement_seconds_per_session, 0.1
    )
  )

knitr::kable(
  landing_status,
  caption = "Identified versus unidentified landing-page sessions"
)
```

GA4 assigned **`r comma(not_set_sessions)` sessions
(`r percent(not_set_share, 0.1)`)** to `(not set)`. This means GA4 recorded the
visit but could not determine its first page. Average engagement time for this
group was only **`r number(landing$average_engagement_time_per_session[
landing$landing_page_original == '(not set)'][1], 0.1)` seconds**. Consent-mode
behaviour, session tracking and redirects are possible causes to investigate;
the table does not identify which cause applies.

The acquisition and landing-page exports differ by **`r comma(session_difference)`
sessions (`r percent(abs(session_difference / total_sessions), 0.1)`)**. The
aggregate results remain useful, but the difference should be treated as a
reporting limitation.

The practical consequence is that landing-page analysis is incomplete:

- `r percent(not_set_share, 0.1)` of landing-page sessions cannot be assigned to
  a specific first page;
- the visible landing-page ranking omits those sessions and may understate some
  real entry pages;
- the report cannot determine which content or campaign generated those visits;
- the very short engagement time for this group may affect overall engagement
  interpretation.

This is a measurement limitation, not evidence that `(not set)` caused visitors
to engage less.

Active-user and new-user values in the identified group are sums across landing
page rows and may contain the same user more than once. They are included for
context, not as a site-wide unique-user total.

## Why the account pages are grouped

**Purpose in this report:** prevent hundreds of ID-specific account URLs from
creating a misleading page ranking. This is a reporting treatment, not a finding
that `/my-account` performs well or poorly.

Many `/my-account` URLs contain a different order number, membership ID,
subscription ID, group ID or joining token. GA4 therefore treats URLs such as
`/my-account/view-order/123` and `/my-account/view-order/456` as separate rows,
even though both represent the same **View order** function. Reporting every ID
separately would make the table long, fragment the activity of one function and
provide little value to a non-technical reader.

Preparation found **`r comma(templated_rows)` original URL rows** of this type
and grouped them into **`r n_distinct(pages$page_path_template[
pages$page_path_was_templated])` functional page groups**. Original URLs remain
available in the clean data; the summary shows the grouped functions so readers
can understand account activity without reading hundreds of ID-specific rows.

```{r}
dynamic_path_overview <- tibble(
  measure = c(
    "Original URL rows containing an ID or token",
    "Different original account URLs",
    "Functional page groups shown in the report",
    "Views across those original URLs"
  ),
  result = comma(c(
    templated_rows,
    n_distinct(pages$page_path_original[pages$page_path_was_templated]),
    n_distinct(pages$page_path_template[pages$page_path_was_templated]),
    sum(pages$views[pages$page_path_was_templated])
  ))
)
knitr::kable(
  dynamic_path_overview,
  caption = "Why hundreds of account URLs become a short summary",
  align = c("l", "r")
)
```

```{r}
pages |>
  filter(page_path_was_templated) |>
  group_by(page_path_template) |>
  summarise(
    original_path_rows = n(),
    views = sum(views),
    events = sum(event_count),
    .groups = "drop"
  ) |>
  arrange(desc(views)) |>
  mutate(views = comma(views), events = comma(events)) |>
  slice_head(n = 15) |>
  knitr::kable(
    caption = "Account activity grouped by page function",
    align = c("l", "r", "r", "r")
  )
```

Active users are deliberately not summed in this table because the same user
may visit more than one original dynamic path.

## Pages to investigate, not pages judged as poor

**`r comma(nrow(review_pages))` pages** met the screening rule of at least 100
views and under 10 seconds average engagement time per active user. These may be
redirects, utility pages, tracking gaps, automated traffic or genuine content
problems; they are review candidates, not automatic failures.

The rule narrows thousands of page rows to a manageable review list. It does not
score page quality. A login page, redirect or file URL may legitimately require
very little engagement time, while a long article with the same result may need
content or tracking investigation.

```{r}
review_pages |>
  transmute(page = page_path_standardised, views = comma(views),
    active_users = comma(active_users),
    `average engagement seconds per active user` = number(
      average_engagement_time_per_active_user, 0.1
    )) |>
  slice_head(n = 15) |>
  knitr::kable(
    caption = "Pages meeting the high-view, low-engagement screening rule",
    align = c("l", "r", "r", "r")
  )
```

# What visitors did and whether outcomes were tracked

::: {.callout-important title="Main message: activity is visible, business outcomes are not"}
GA4 records page views, scrolling, clicks, downloads and form interactions, but
all four supplied reports show **`r comma(reported_key_events)` Key Events**.
Therefore, this report can describe what visitors did on the website but cannot
show how many completed a registration, membership action or purchase. Zero here
means **not measurable in these exports**, not necessarily zero real outcomes.
:::

```{r}
knitr::kable(
  key_event_reconciliation |>
    mutate(key_events = comma(key_events)),
  caption = "Checking whether all GA4 reports agree on Key Events",
  align = c("l", "r")
)
```

The export contains **`r comma(page_view$event_count)` page views**,
**`r comma(events$event_count[events$event_name == 'scroll'])` scrolls**,
**`r comma(events$event_count[events$event_name == 'click'])` clicks** and
**`r comma(events$event_count[events$event_name == 'file_download'])` file
downloads**. Under standard GA4 Enhanced Measurement, `scroll` fires only the
first time 90% vertical depth becomes visible on each page—not every time a user
moves the scroll wheel. The observed scroll count is approximately
**`r percent(scroll_to_view_ratio, 0.1)` of page views**. This is not a true
reader-completion rate because page length, automatic visibility, consent and
implementation can affect the event, but it explains why scroll events are
substantially fewer than page views.

The export also contains **`r comma(events$event_count[events$event_name ==
'form_start'])` form starts** and **`r comma(events$event_count[events$event_name
== 'form_submit'])` form submissions**.

These form names suggest intent but do not identify the form or business
outcome. More importantly, supplied Key Events total **`r reported_key_events`**.
Conversion rates by channel, landing page and content page are therefore not
available. STAV should agree the outcomes that matter—such as completed event
registration, membership application or renewal, resource purchase and contact
enquiry—then configure and test them as Key Events. Stable `form_name` or
`form_id` parameters should distinguish generic form events.

```{r}
events |>
  arrange(desc(event_count)) |>
  transmute(event_name, category = event_category,
    event_count = comma(event_count), users = comma(total_users),
    events_per_active_user = number(event_count_per_active_user, 0.01)) |>
  knitr::kable(align = c("l", "l", "r", "r", "r"))
```

# Countries that need a closer look

**`r nrow(flagged_countries)` high-volume country rows** meet the review rule:
at least 500 active users and either engagement below 15% or average engagement
time below 10 seconds. These flags are not proof of automated or irrelevant
traffic. Country-level users can also overlap. Hostname, source, medium,
landing-page and device evidence is required before filtering decisions.

The rule is an analyst-created screening tool, not a Google threshold. The
500-user minimum prevents tiny country rows from being flagged because of a few
short visits. The engagement thresholds identify rows that are both large
enough to affect overall results and unusually weak on at least one engagement
measure. A flag means **investigate**, not **exclude**.

```{r}
review_rule <- tribble(
  ~component, ~condition, ~purpose,
  "Volume gate", "Active users >= 500",
    "Focus review on country rows large enough to affect the analysis",
  "Engagement signal 1", "Engagement rate < 15%",
    "Identify country rows where few sessions qualify as engaged",
  "Engagement signal 2", "Average engagement time/user < 10 seconds",
    "Identify country rows with very short measured attention",
  "Final flag", "Volume gate AND either engagement signal",
    "Create a review queue; do not delete or classify traffic automatically"
)
knitr::kable(review_rule, caption = "How the geographic review rule works")
```

```{r}
flagged_countries |>
  transmute(country, active_users = comma(active_users), new_users = comma(new_users),
    engagement_rate = percent(engagement_rate, 0.1),
    `average engagement seconds per active user` = number(
      average_engagement_time_per_active_user, 0.1
    ),
    events = comma(event_count)) |>
  knitr::kable(align = c("l", "r", "r", "r", "r", "r"))
```

# What this report can and cannot answer

```{r}
coverage <- tribble(
  ~question, ~status, ~reason,
  "How much aggregate traffic and content activity occurred?", "Supported",
    "Sessions, views and events cover the full period",
  "Which default channels and pages received the most activity?", "Supported",
    "Channel, landing-page and page-path dimensions are supplied",
  "How did traffic change by month?", "Not supported", "No date/month dimension",
  "Which source, medium or campaign drove results?", "Not supported",
    "Only default channel group is present",
  "How do New and Returning users compare?", "Not supported",
    "No explicit Returning-user dimension",
  "Which channels or pages generated conversions?", "Not supported",
    "All supplied Key Event totals are zero",
  "Where do users exit or navigate next?", "Not supported",
    "Navigation-path and exit dimensions are absent"
)
knitr::kable(coverage, align = c("l", "l", "l"))
```

# Scope boundary

The pipeline changes only analysis-layer files. Raw exports remain unchanged.
Tracking changes and traffic filters require implementation and validation by
STAV's authorised administrators.
