---
title: "STAV Website Analytics Report"
subtitle: "Google Analytics 4 Analysis"
author: "Feng Zhu"
date: "2026-09-30"

format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    embed-resources: true

execute:
  echo: false
  warning: false
  message: false

editor: visual
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(janitor)
library(lubridate)
```

```{r}
#| label: load-data
#| include: false

monthly <- read_csv(
  "../data/GA4/traffic_channel_monthly_20250701_20260630.csv",
  skip = 9,
  show_col_types = FALSE
) %>%
  clean_names()

channels <- read_csv(
  "../data/GA4/traffic_channel_20250701_20260630.csv",
  skip = 9,
  show_col_types = FALSE
) %>%
  clean_names()

source_medium <- read_csv(
  "../data/GA4/traffic_source_medium_20250701_20260630.csv",
  skip = 9,
  show_col_types = FALSE
) %>%
  clean_names()

landing <- read_csv(
  "../data/GA4/landing_page_20250701_20260630.csv",
  skip = 9,
  show_col_types = FALSE
) %>%
  clean_names()

pages <- read_csv(
  "../data/GA4/pages_screens_page_path_20250701_20260630.csv",
  skip = 9,
  show_col_types = FALSE
) %>%
  clean_names()

monthly_pages <- read_csv(
  "../data/GA4/pages_screens_page_path_month_20250701_20260630.csv",
  skip = 9,
  show_col_types = FALSE
) %>%
  clean_names()

device <- read_csv(
  "../data/GA4/device_category_20250701_20260630.csv",
  skip = 9,
  show_col_types = FALSE
) %>%
  clean_names()

country <- read_csv(
  "../data/GA4/country_20250701_20260630.csv",
  skip = 9,
  show_col_types = FALSE
) %>%
  clean_names()

events <- read_csv(
  "../data/GA4/events_20250701_20260630.csv",
  skip = 9,
  show_col_types = FALSE
) %>%
  clean_names()
```

# Executive Summary

This report analyses STAV website activity using Google Analytics 4 (GA4) data for the period from 1 July 2025 to 30 June 2026. The analysis examines how users reached the website, which content attracted the most activity, the devices and geographic locations associated with website use, and the interactions recorded across the site.

The analysis identified six key findings:

1.  **Most website traffic came from Direct and Organic Search.** Direct accounted for approximately 50.6% of sessions, while Organic Search accounted for 43.5%. Together, these two channels generated around 94% of website sessions. Organic Search also recorded a higher engagement rate (59.3%) than Direct traffic (29.2%), showing that a larger proportion of search sessions met GA4's engagement criteria.

2.  **Science Talent Search (STS) was a central component of website activity.** Science Talent Search content accounted for approximately 31.7% of landing-page sessions and 35.3% of all page views. This means it was important both as a way for visitors to enter the website and as content they viewed while using the site, making it a key area for STAV to maintain and monitor.

3.  **Website attention was concentrated around a relatively small proportion of content.** Website activity was concentrated on a relatively small number of pages. The website has many different pages, but the 100 most-viewed pages generated approximately 87.2% of all page views. This means STAV can focus more attention on maintaining its most frequently viewed pages rather than reviewing every page equally.

4.  **Website usage was strongly desktop-oriented, while mobile users still showed meaningful engagement.** Desktop represented approximately 83.4% of device-level Active Users, compared with 15.9% for mobile. However, mobile recorded a higher engagement rate of 50.4%, compared with 42.0% for desktop. This suggests that mobile remains an important part of the website experience despite its smaller user volume.

5.  **Meaningful website activity was strongly concentrated in Australia.** Australia represented 64.9% of country-level Active Users and generated 85.5% of recorded events. Several international locations recorded relatively high user counts but substantially weaker engagement signals.

6.  **Visitors interacted with the website in several ways beyond simply viewing pages.** GA4 recorded substantial activity including scrolling, form interactions, clicks and file downloads. However, the current tracking does not provide enough detail to identify which specific forms, files or resources generated these interactions. This limits STAV's ability to understand which website activities are most important to users.

Based on these findings, the report recommends prioritising high-demand content such as Science Talent Search, maintaining STAV's Organic Search visibility, regularly testing important mobile website tasks, improving the detail captured for forms and file downloads, and tracking the website actions that matter most to STAV.

# Project Objective and Scope

The purpose of this analysis is to help STAV understand how people use its website, where website traffic comes from, which content receives the most attention, and how website measurement could be improved in the future.

The analysis focuses on website activity recorded in Google Analytics 4 (GA4) between **1 July 2025 and 30 June 2026**. It examines overall traffic patterns, acquisition channels and sources, landing pages, website content, device usage, geographic distribution and tracked user behaviour.

The analysis addresses four main questions:

1.  How do users reach the STAV website, and how does website traffic change over time?
2.  Which website content and entry points attract the most activity?
3.  What do the available device and geographic data indicate about the website audience?
4.  What user interactions are currently captured by GA4, and where are there opportunities to improve measurement?

This report focuses specifically on the **GA4 website analytics component** of the internship project.

The analysis is intended to support practical decision-making rather than evaluate website performance against an external benchmark. Findings are therefore interpreted using patterns observed within STAV's own GA4 data.

# Methodology

## Data Source

Website analytics data were obtained from STAV's Google Analytics 4 (GA4) property. The reporting period was **1 July 2025 to 30 June 2026**, providing a full year of website activity for analysis.

Separate GA4 reports were exported for traffic acquisition, monthly traffic trends, source/medium, landing pages, pages and screens, device categories, countries and events. These exports provided different views of website acquisition, content usage, audience characteristics and tracked user interactions.

## Data Preparation

The exported GA4 data were cleaned and analysed in R. The data were checked for missing values, duplicate records and inconsistencies before analysis. Monthly data were organised chronologically from July 2025 to June 2026. GA4 report totals were also compared with the exported data as a quality check. Where small differences occurred between reports, the original GA4 values were retained rather than manually adjusted.

## Content Classification

To make the website data easier to understand, individual pages were grouped into broader content categories such as Science Talent Search, Events & Conferences, Resources & Publications, Shop, Membership and Account.

These categories were created specifically for this analysis and are not standard GA4 categories. The original page-level data were retained so that the classifications could be checked where needed.

## Analysis Approach

The analysis examined website traffic over time, how visitors reached the website, which content they viewed, the devices and countries associated with website activity, and the interactions recorded by GA4.

Some GA4 metrics require careful interpretation. In particular, Active Users cannot simply be added across pages, devices or countries, because the same person may appear in more than one row. For example, someone who visits both the homepage and a Science Talent Search page may be counted as an Active User for both pages. Therefore, the GA4 report-level total of 55,599 Active Users is used when describing the overall website audience.

The analysis identifies patterns in the available data, but does not assume that a particular STAV activity caused an increase or decrease in website traffic unless supporting evidence is available.

# Website Traffic and Acquisition

## Overall Traffic Trend

```{r}
#| label: fig-monthly-traffic
#| fig-cap: "STAV website traffic by month, July 2025 to June 2026"
#| fig-width: 9
#| fig-height: 5

monthly_total <- monthly %>%
  mutate(
    month_num = as.integer(month),
    year = if_else(month_num >= 7, 2025L, 2026L),
    date = ymd(sprintf("%d-%02d-01", year, month_num))
  ) %>%
  group_by(date) %>%
  summarise(
    sessions = sum(sessions, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(
  monthly_total,
  aes(x = date, y = sessions)
) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  
  # Show session values
  geom_text(
    aes(label = scales::comma(sessions)),
    vjust = -0.8,
    size = 3.5
  ) +
  
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b\n%Y"
  ) +
  scale_y_continuous(
    labels = scales::comma,
    expand = expansion(mult = c(0.05, 0.15))
  ) +
  labs(
    title = "Website Traffic by Month",
    subtitle = "STAV, July 2025 – June 2026",
    x = NULL,
    y = "Website sessions"
  ) +
  theme_minimal()
```

*In GA4, a session represents a period in which a user interacts with the website. One person may generate multiple sessions if they return to the website at different times.*

**What this means for STAV:** Website activity changed substantially throughout the year. Traffic was lowest in September 2025, increased strongly in October, and was generally higher between February and June 2026, reaching its highest level in May.

These changes may reflect the timing of STAV activities, events, registrations, resource releases or other communications that encouraged people to visit the website. Lower-traffic periods may similarly reflect periods with fewer time-sensitive activities or lower demand for website content. However, the available GA4 data does not identify the specific cause of these monthly changes, so these factors should be treated as possible explanations rather than confirmed causes.

## Acquisition Channels

Website traffic was dominated by two acquisition channels: **Direct** and **Organic Search**. Direct generated 46,708 sessions (approximately 50.6% of annual channel sessions), while Organic Search generated 40,092 sessions (43.5%). Referral traffic contributed a further 4,190 sessions (4.5%), while all other channels represented only a small proportion of recorded traffic.

```{r}
#| label: fig-acquisition-channels
#| fig-cap: "How visitors reached the STAV website"
#| fig-width: 9
#| fig-height: 5

channel_summary <- channels %>%
  mutate(
    session_share = sessions / sum(sessions)
  ) %>%
  arrange(sessions)

ggplot(
  channel_summary,
  aes(
    x = reorder(
      session_primary_channel_group_default_channel_group,
      sessions
    ),
    y = sessions
  )
) +
  geom_col() +
  geom_text(
    aes(
      label = paste0(
        scales::comma(sessions),
        " (",
        scales::percent(session_share, accuracy = 0.1),
        ")"
      )
    ),
    hjust = -0.1,
    size = 3.5
  ) +
  coord_flip() +
  scale_y_continuous(
    labels = scales::comma,
    expand = expansion(mult = c(0, 0.25))
  ) +
  labs(
    title = "How Visitors Reached the STAV Website",
    subtitle = "STAV, July 2025 – June 2026",
    x = NULL,
    y = "Website sessions"
  ) +
  theme_minimal()
```

**How to read this chart:**\
**Direct** means GA4 could not identify another referring source for the session. **Organic Search** means visitors arrived through unpaid search-engine results, such as Google or Bing. **Referral** means visitors arrived by following a link from another website.

**What this means for STAV:** Direct and Organic Search dominated website acquisition. Direct generated 46,708 sessions (50.6%), while Organic Search generated 40,092 sessions (43.5%). Together, they accounted for approximately 94% of annual channel sessions.

Organic Search also showed stronger engagement signals. Its engagement rate was 59.3%, compared with 29.2% for Direct traffic. This means Organic Search sessions were more frequently classified by GA4 as engaged sessions; it does not necessarily mean these visitors were more valuable to STAV.

## Traffic Sources

The Source / Medium report provides a more detailed view of where website traffic came from. While the previous chart groups traffic into broad channels, this analysis identifies the specific sources within those channels.

```{r}
#| label: fig-top-traffic-sources
#| fig-cap: "Top three traffic sources to the STAV website"
#| fig-width: 9
#| fig-height: 4.5

source_medium_total <- sum(source_medium$sessions)

top_sources <- source_medium %>%
  filter(session_source_medium != "(not set)") %>%
  mutate(
    session_share = sessions / source_medium_total
  ) %>%
  slice_max(
    order_by = sessions,
    n = 3,
    with_ties = FALSE
  )

ggplot(
  top_sources,
  aes(
    x = reorder(session_source_medium, sessions),
    y = sessions
  )
) +
  geom_col() +
  geom_text(
    aes(
      label = paste0(
        scales::comma(sessions),
        " (",
        scales::percent(session_share, accuracy = 0.1),
        ")"
      )
    ),
    hjust = -0.1,
    size = 4
  ) +
  coord_flip() +
  scale_y_continuous(
    labels = scales::comma,
    expand = expansion(mult = c(0, 0.25))
  ) +
  labs(
    title = "Top Three Traffic Sources",
    subtitle = "STAV, July 2025 – June 2026",
    x = NULL,
    y = "Website sessions"
  ) +
  theme_minimal()
```

**What does "(direct) / (none)" mean?**\
In GA4, "(direct) / (none)" means that GA4 did not identify another source or medium for the session. This can include users who accessed the website directly, such as through a bookmark or by entering the website address, but it can also include visits where referral information was unavailable. It should therefore not be interpreted as only users manually typing the STAV website address.

**What this means for STAV:** The previous chart showed that Organic Search was one of STAV's two dominant acquisition channels. Looking more closely at the underlying sources shows that most of this search traffic came from Google (35,636 sessions) and Bing (3,805 sessions). Together with Direct traffic, these three sources accounted for approximately 93.6% of source-level sessions.

