---
title: "CRM data preparation"
subtitle: "Import cleaning standardisation and reusable validation flags"
author: "Jordan"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    code-fold: true
    embed-resources: true
execute:
  warning: false
  message: false
---

# Purpose

This report creates the reusable cleaned CRM contact dataset used by all later
CRM analyses. It imports every source field, standardises values, adds technical
validation flags and saves record-level data in the restricted output folder.

This report does not interpret business results. Record counts, duplicate
analysis, completeness, school groups, membership analysis and contact activity
are handled in later reports.

# 1 Setup and source contract

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(here)

options(scipen = 999)

here::i_am("reports/CRM/01_crm_cleaning.qmd")

crm_file <- here("data", "raw", "CRM", "CRM20260729.csv")
extract_date <- as.Date("2026-07-29")

stopifnot(file.exists(crm_file))
```

# 2 Import all source fields

All fields are imported as text so that identifiers, phone numbers and mixed
date formats are not altered by automatic type guessing.

```{r}
crm_raw <- suppressWarnings(
  read_csv(
    crm_file,
    col_types = cols(.default = col_character()),
    trim_ws = FALSE,
    name_repair = "minimal",
    show_col_types = FALSE,
    progress = FALSE,
    lazy = FALSE
  )
)

source_record_count <- nrow(crm_raw)
source_field_count <- ncol(crm_raw)
source_column_names <- names(crm_raw)
source_exact_duplicate_rows <- sum(duplicated(crm_raw))

crm_parse_problems <- problems(crm_raw)
problem_record_numbers <- sort(unique(crm_parse_problems$row - 1L))

crm_import_validation <- tibble(
  check = c(
    "Source records",
    "Source fields",
    "Parser problem entries",
    "Affected source records",
    "Exact duplicate source rows"
  ),
  result = c(
    source_record_count,
    source_field_count,
    nrow(crm_parse_problems),
    length(problem_record_numbers),
    source_exact_duplicate_rows
  )
)

crm_import_validation
```

The table is based on the raw 71-field CRM import. The rows document the size and structural condition
of the raw import before cleaning. Parser entries and affected records are not
the same measure because one record can generate more than one parser entry.
Exact duplicates require all 71 source fields to match.

# 3 Standardise structure and missing values

## 3.1 Standardise column names and add audit fields

```{r}
crm_clean <- crm_raw |>
  clean_names() |>
  mutate(
    source_record_number = row_number(),
    structure_issue = source_record_number %in% problem_record_numbers,
    .before = 1
  )

source_fields_clean_names <- names(crm_clean)[
  !names(crm_clean) %in% c("source_record_number", "structure_issue")
]

stopifnot(length(source_fields_clean_names) == source_field_count)
```

## 3.2 Clean whitespace and blank strings

```{r}
crm_clean <- crm_clean |>
  mutate(
    across(
      all_of(source_fields_clean_names),
      ~ .x |> str_squish() |> na_if("")
    )
  )
```

## 3.3 Quarantine structurally unreliable values

For parser-affected rows, values from `notes` onward are not reliable. Those
values are set to missing, while earlier contact fields and the source row are
retained for traceability.

```{r}
post_notes_fields <- source_fields_clean_names[
  match("notes", source_fields_clean_names):
    match("company_department", source_fields_clean_names)
]

crm_clean <- crm_clean |>
  mutate(
    across(
      all_of(post_notes_fields),
      ~ if_else(structure_issue, NA_character_, .x)
    )
  )
```

## 3.4 Convert known missing-value placeholders

`No` is converted to missing only in fields where it represents information
that was not supplied. Genuine yes/no response fields are not changed.

```{r}
no_as_missing_fields <- intersect(
  c(
    "birthday", "mobile_phone", "primary_phone",
    "primary_phone_extension", "street_address_1", "street_address_2",
    "city", "postal_zip", "region", "country", "time_zone",
    "ip_address", "lead_source", "source_page", "gdpr_consent_date",
    "marketing_consent_date", "notes", "tags",
    "utm_campaign", "utm_content", "utm_medium", "utm_term", "utm_source",
    "unsub_date", "unsub_reason", "unsub_feedback",
    "membership_status", "membership_plan", "subscription_start_date",
    "subscription_end_date", "date_of_cancellation",
    "reason_for_cancellation", "if_other", "date_retired", "year_level",
    "subjects", "job_title_position",
    "do_you_hold_an_identified_leadership_role_other",
    "your_role", "last_sponsorship_purchase", "sponsorships_details_sections",
    "sponsorship_details_divisions", "supporting_sponsorships",
    "company_name", "company_website", "company_address", "company_phone",
    "company_phone_extension", "job_title", "company_department"
  ),
  names(crm_clean)
)

no_to_na <- function(x) {
  if_else(
    !is.na(x) & str_to_lower(str_squish(x)) == "no",
    NA_character_,
    x
  )
}

crm_clean <- crm_clean |>
  mutate(across(all_of(no_as_missing_fields), no_to_na))
```

# 4 Create reusable comparison keys

Keys are added for matching and validation without overwriting original CRM
values.

```{r}
normalise_name_part <- function(x) {
  x |>
    str_to_lower() |>
    str_squish() |>
    str_replace_all("[^[:alnum:]]", "")
}

normalise_school_name <- function(x) {
  x |>
    str_to_lower() |>
    str_squish() |>
    str_replace_all("[^[:alnum:] ]", "") |>
    str_squish()
}

crm_clean <- crm_clean |>
  mutate(
    email_key = email |> str_to_lower() |> str_squish(),
    first_name_key = normalise_name_part(first_name),
    last_name_key = normalise_name_part(last_name),
    name_key = case_when(
      !is.na(first_name_key) & !is.na(last_name_key) ~
        str_c(first_name_key, last_name_key, sep = "|"),
      TRUE ~ NA_character_
    ),
    supplied_full_name_key = normalise_name_part(full_name),
    constructed_full_name_key = case_when(
      !is.na(first_name_key) & !is.na(last_name_key) ~
        str_c(first_name_key, last_name_key),
      TRUE ~ NA_character_
    ),
    school_name_key = normalise_school_name(company_name),
    tags_key = tags |> str_to_lower() |> str_squish()
  )
```

# 5 Standardise reusable contact and location fields

The source does not contain a field literally named `suburb`. `city` is retained
as `suburb_candidate` for later business confirmation. A preferred phone field
uses mobile first and primary phone second without discarding either source
field.

```{r}
normalise_phone <- function(x) {
  digits <- str_replace_all(x, "[^0-9+]", "")
  na_if(digits, "")
}

normalise_region <- function(x) {
  key <- x |> str_to_lower() |> str_squish()

  case_when(
    key %in% c("vic", "victoria") ~ "VIC",
    key %in% c("nsw", "new south wales") ~ "NSW",
    key %in% c("qld", "queensland") ~ "QLD",
    key %in% c("act", "australian capital territory") ~ "ACT",
    key %in% c("sa", "south australia") ~ "SA",
    key %in% c("wa", "western australia") ~ "WA",
    key %in% c("tas", "tasmania") ~ "TAS",
    key %in% c("nt", "northern territory") ~ "NT",
    key == "metro" ~ "Metro",
    key == "country" ~ "Country",
    is.na(key) ~ NA_character_,
    TRUE ~ str_to_title(key)
  )
}

crm_clean <- crm_clean |>
  mutate(
    mobile_phone_clean = normalise_phone(mobile_phone),
    primary_phone_clean = normalise_phone(primary_phone),
    preferred_phone = coalesce(mobile_phone_clean, primary_phone_clean),
    suburb_candidate = city,
    region_clean = normalise_region(region),
    postal_zip_clean = postal_zip |> str_replace_all("[^0-9]", "") |> na_if(""),
    lead_source_clean = lead_source |> str_to_lower() |> str_squish()
  )
```

# 6 Parse reusable date fields

```{r}
parse_crm_date <- function(x) {
  as.Date(
    suppressWarnings(
      parse_date_time(
        x,
        orders = c(
          "ymd HMS", "ymd HM", "ymd",
          "dmy HMS", "dmy HM", "dmy",
          "mdy HMS", "mdy HM", "mdy",
          "B d Y IMS p", "B d Y IM p"
        ),
        tz = "Australia/Melbourne",
        quiet = TRUE
      )
    )
  )
}

date_fields <- intersect(
  c(
    "date_created", "date_optin_status_changed", "gdpr_consent_date",
    "marketing_consent_date", "unsub_date", "subscription_start_date",
    "subscription_end_date", "date_of_cancellation", "date_retired",
    "last_sponsorship_purchase"
  ),
  names(crm_clean)
)

crm_clean <- crm_clean |>
  mutate(
    across(
      all_of(date_fields),
      parse_crm_date,
      .names = "{.col}_clean"
    )
  )

crm_date_parse_validation <- map_dfr(date_fields, function(field_name) {
  raw_values <- crm_clean[[field_name]]
  clean_values <- crm_clean[[paste0(field_name, "_clean")]]

  tibble(
    field = field_name,
    non_missing_raw = sum(!is.na(raw_values)),
    successfully_parsed = sum(!is.na(clean_values)),
    parse_failures = sum(!is.na(raw_values) & is.na(clean_values)),
    earliest_date = if (all(is.na(clean_values))) {
      as.Date(NA)
    } else {
      min(clean_values, na.rm = TRUE)
    },
    latest_date = if (all(is.na(clean_values))) {
      as.Date(NA)
    } else {
      max(clean_values, na.rm = TRUE)
    }
  )
})

crm_date_parse_validation
```

The table is based on each available source date field and its corresponding
`_clean` field. Each row compares populated source date values with
successfully parsed clean dates. A parse failure means a non-missing source
value could not be converted using the supported formats; it does not by itself
show whether the source date is factually correct.

# 7 Standardise membership source values

This section standardises source values for reuse. Membership interpretation
and reporting are performed in `04_crm_membership_analysis.qmd`.

```{r}
normalise_membership_status <- function(x) {
  key <- x |>
    str_to_lower() |>
    str_squish() |>
    str_replace_all("[ _]", "-")

  case_when(
    key %in% c(
      "active", "on-hold", "expired", "cancelled", "pending-cancel",
      "paused", "pending"
    ) ~ key,
    is.na(key) ~ NA_character_,
    TRUE ~ "review-unmapped"
  )
}

normalise_membership_plan <- function(x) {
  x |>
    str_replace_all("[–—]", "-") |>
    str_replace_all("\\s*-\\s*", " - ") |>
    str_squish()
}

crm_clean <- crm_clean |>
  mutate(
    membership_status_clean = normalise_membership_status(membership_status),
    membership_plan_clean = normalise_membership_plan(membership_plan)
  )
```

## 7.2 Classify membership tags

The three requested membership tags are converted into a reusable category.
The classification follows the agreed priority order: Active, On-hold, then
Expired. Separate flags are retained so later analysis can identify contacts
carrying more than one of these tags.

```{r}
crm_clean <- crm_clean |>
  mutate(
    membership_tag_active = str_detect(
      coalesce(tags, ""),
      regex("\\[MEMBERSHIP\\] ACTIVE", ignore_case = TRUE)
    ),
    membership_tag_on_hold = str_detect(
      coalesce(tags, ""),
      regex("\\[MEMBERSHIP\\] ON-HOLD", ignore_case = TRUE)
    ),
    membership_tag_expired = str_detect(
      coalesce(tags, ""),
      regex("\\[MEMBERSHIP\\] EXPIRED", ignore_case = TRUE)
    ),
    membership_tag_status = case_when(
      membership_tag_active ~ "active",
      membership_tag_on_hold ~ "on-hold",
      membership_tag_expired ~ "expired",
      TRUE ~ NA_character_
    ),
    membership_tag_category_count =
      as.integer(membership_tag_active) +
      as.integer(membership_tag_on_hold) +
      as.integer(membership_tag_expired),
    multiple_membership_tag_categories = membership_tag_category_count > 1L
  )
```

## 7.3 Create the analysis-ready variable set

All 71 source fields remain in `crm_clean` for audit purposes. The downstream
analysis dataset is restricted to the 25 agreed business variables, plus audit,
standardised and validation fields created by this report.

```{r}
selected_business_fields <- c(
  "id", "full_name", "first_name", "last_name", "email",
  "optin_status", "user_id", "owner_id", "date_created",
  "date_optin_status_changed", "city", "postal_zip", "region",
  "lead_source", "tags", "membership_status", "membership_plan",
  "subscription_start_date", "subscription_end_date", "year_level",
  "subjects", "job_title_position",
  "do_you_hold_an_identified_leadership_role",
  "do_you_hold_an_identified_leadership_role_other", "company_name"
)

stopifnot(length(selected_business_fields) == 25L)
stopifnot(all(selected_business_fields %in% names(crm_clean)))
```

# 8 Create technical validation flags

Flags identify records for later analysis or manual review. They do not remove
records or make final business classifications.

```{r}
flag_all_duplicates <- function(x) {
  !is.na(x) & (duplicated(x) | duplicated(x, fromLast = TRUE))
}

crm_clean <- crm_clean |>
  mutate(
    duplicate_crm_id_candidate = flag_all_duplicates(id),
    duplicate_email_candidate = flag_all_duplicates(email_key),
    duplicate_name_candidate = flag_all_duplicates(name_key),
    email_format_valid = case_when(
      is.na(email_key) ~ NA,
      TRUE ~ str_detect(
        email_key,
        "^[^[:space:]@]+@[^[:space:]@]+\\.[^[:space:]@]+$"
      )
    ),
    full_name_component_mismatch =
      !is.na(supplied_full_name_key) &
      !is.na(constructed_full_name_key) &
      supplied_full_name_key != constructed_full_name_key,
    australian_postcode_format_valid = case_when(
      is.na(postal_zip_clean) ~ NA,
      TRUE ~ str_detect(postal_zip_clean, "^[0-9]{4}$")
    ),
    membership_end_before_start =
      !is.na(subscription_start_date_clean) &
      !is.na(subscription_end_date_clean) &
      subscription_end_date_clean < subscription_start_date_clean,
    active_membership_with_past_end_date =
      membership_status_clean == "active" &
      !is.na(subscription_end_date_clean) &
      subscription_end_date_clean < extract_date
  )
```

# 9 Verify the output contract

```{r}
required_source_fields <- c(
  "id", "full_name", "first_name", "last_name", "email",
  "mobile_phone", "primary_phone", "city", "postal_zip", "region",
  "tags", "membership_status", "membership_plan",
  "subscription_start_date", "subscription_end_date", "company_name"
)

required_derived_fields <- c(
  "source_record_number", "structure_issue", "email_key", "name_key",
  "school_name_key", "preferred_phone", "suburb_candidate", "region_clean",
  "postal_zip_clean", "membership_status_clean", "membership_plan_clean",
  "membership_tag_active", "membership_tag_on_hold",
  "membership_tag_expired", "membership_tag_status",
  "membership_tag_category_count",
  "multiple_membership_tag_categories",
  "duplicate_crm_id_candidate", "duplicate_email_candidate",
  "duplicate_name_candidate", "email_format_valid",
  "full_name_component_mismatch",
  "australian_postcode_format_valid"
)

stopifnot(nrow(crm_clean) == source_record_count)
stopifnot(all(required_source_fields %in% names(crm_clean)))
stopifnot(all(required_derived_fields %in% names(crm_clean)))
stopifnot(!anyDuplicated(crm_clean$source_record_number))

crm_analysis <- crm_clean |>
  select(
    source_record_number,
    structure_issue,
    all_of(selected_business_fields),
    all_of(required_derived_fields[!required_derived_fields %in% c(
      "source_record_number", "structure_issue"
    )]),
    ends_with("_clean"),
    membership_end_before_start,
    active_membership_with_past_end_date
  ) |>
  select(!any_of(c("mobile_phone_clean", "primary_phone_clean")))

stopifnot(nrow(crm_analysis) == source_record_count)
stopifnot(all(selected_business_fields %in% names(crm_analysis)))

crm_output_validation <- tibble(
  check = c(
    "Output records equal source records",
    "All source fields retained",
    "Required source fields available",
    "Required derived fields available",
    "Analysis dataset contains 25 selected source fields",
    "Source record numbers unique"
  ),
  passed = c(
    nrow(crm_clean) == source_record_count,
    all(source_fields_clean_names %in% names(crm_clean)),
    all(required_source_fields %in% names(crm_clean)),
    all(required_derived_fields %in% names(crm_clean)),
    all(selected_business_fields %in% names(crm_analysis)),
    !anyDuplicated(crm_clean$source_record_number)
  )
)

crm_output_validation
```

The table is based on the source row count, source field list, selected 25-field
contract and derived-field contract. Every row is a pipeline integrity test. `TRUE` means
the clean output retained the expected records and fields and preserved unique
source row identifiers. These checks validate the pipeline structure, not the
business accuracy of every CRM value.

# 10 Save pipeline outputs

## 10.1 Output locations

```{r}
output_dir <- here("data", "clean", "CRM")
restricted_dir <- file.path(output_dir, "restricted")

dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(restricted_dir, recursive = TRUE, showWarnings = FALSE)
```

## 10.2 Record-level restricted outputs

```{r}
crm_review_flags_internal <- crm_clean |>
  filter(
    structure_issue |
      duplicate_crm_id_candidate |
      duplicate_email_candidate |
      duplicate_name_candidate |
      coalesce(!email_format_valid, FALSE) |
      coalesce(!australian_postcode_format_valid, FALSE) |
      coalesce(membership_end_before_start, FALSE) |
      coalesce(active_membership_with_past_end_date, FALSE)
  )

write_csv(
  crm_clean,
  file.path(restricted_dir, "crm_contacts_clean_internal.csv"),
  na = ""
)

write_csv(
  crm_analysis,
  file.path(restricted_dir, "crm_contacts_analysis_internal.csv"),
  na = ""
)

write_csv(
  crm_review_flags_internal,
  file.path(restricted_dir, "crm_review_flags_internal.csv"),
  na = ""
)
```

## 10.3 Non-identifiable technical outputs

```{r}
crm_cleaning_log <- tribble(
  ~step, ~area, ~rule,
  1L, "Import", "Import all source variables as character values",
  2L, "Audit", "Retain source row number and parser issue flag",
  3L, "Missing values", "Standardise whitespace blanks and selected No placeholders",
  4L, "Structure", "Quarantine unreliable post-notes values in parser-affected rows",
  5L, "Matching", "Create reusable email name school and tag keys",
  6L, "Contact", "Create clean phone suburb candidate region and postcode fields",
  7L, "Dates", "Parse reusable CRM dates without overwriting source text",
  8L, "Membership", "Standardise membership status plan and requested tag category",
  9L, "Selection", "Create analysis data from the 25 agreed business variables",
  10L, "Validation", "Create technical flags without deleting records",
  11L, "Output", "Separate restricted records from aggregate technical outputs"
)

crm_analysis_variable_contract <- tibble(
  position = seq_along(selected_business_fields),
  source_field = selected_business_fields
)

write_csv(
  crm_import_validation,
  file.path(output_dir, "crm_import_validation.csv"),
  na = ""
)

write_csv(
  crm_date_parse_validation,
  file.path(output_dir, "crm_date_parse_validation.csv"),
  na = ""
)

write_csv(
  crm_output_validation,
  file.path(output_dir, "crm_output_validation.csv"),
  na = ""
)

write_csv(
  crm_cleaning_log,
  file.path(output_dir, "crm_cleaning_log.csv"),
  na = ""
)

write_csv(
  crm_analysis_variable_contract,
  file.path(output_dir, "crm_analysis_variable_contract.csv"),
  na = ""
)
```

# 11 Handoff to analysis reports

The reusable record-level input for all later CRM reports is:

```text
data/clean/CRM/restricted/crm_contacts_analysis_internal.csv
```

The next report, `02_crm_data_quality.qmd`, will calculate and interpret record
counts, duplicates, invalid contact details and business-field completeness.
