---
title: "CRM data quality analysis"
subtitle: "Record volume duplicates validity and completeness"
author: "Jordan"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: true
    code-fold: true
    embed-resources: true
execute:
  warning: false
  message: false
---

# Purpose

This report evaluates the quality of the 25 agreed CRM business variables. It
answers four questions:

1. How many contact records are available?
2. How many duplicate or possible duplicate records are present?
3. Are key identifiers and contact fields technically valid?
4. How complete are the selected variables?

Potential issues are flagged for review rather than deleted. Membership category
reconciliation, school group analysis and active or dormant contact analysis are
handled in later reports.

# 1 Setup and input contract

```{r}
library(tidyverse)
library(lubridate)
library(here)

options(scipen = 999)
here::i_am("reports/CRM/02_crm_data_quality.qmd")

crm_file <- here(
  "data", "clean", "CRM", "restricted",
  "crm_contacts_analysis_internal.csv"
)

variable_contract_file <- here(
  "data", "clean", "CRM", "crm_analysis_variable_contract.csv"
)

import_validation_file <- here(
  "data", "clean", "CRM", "crm_import_validation.csv"
)

stopifnot(file.exists(crm_file))
stopifnot(file.exists(variable_contract_file))
stopifnot(file.exists(import_validation_file))

crm <- read_csv(
  crm_file,
  col_types = cols(.default = col_character()),
  show_col_types = FALSE,
  progress = FALSE
)

variable_contract <- read_csv(
  variable_contract_file,
  show_col_types = FALSE
)

import_validation <- read_csv(
  import_validation_file,
  show_col_types = FALSE
)

exact_duplicate_source_rows <- import_validation |>
  filter(check == "Exact duplicate source rows") |>
  pull(result)

stopifnot(length(exact_duplicate_source_rows) == 1L)

selected_business_fields <- variable_contract$source_field

stopifnot(length(selected_business_fields) == 25L)
stopifnot(all(selected_business_fields %in% names(crm)))
stopifnot(!anyDuplicated(crm$source_record_number))
```

# 2 Record volume

```{r}
crm_record_summary <- tibble(
  metric = c(
    "Contact records",
    "Distinct CRM IDs",
    "Distinct normalised emails",
    "Records with a school or organisation",
    "Records with membership status",
    "Records with membership plan"
  ),
  records = c(
    nrow(crm),
    n_distinct(crm$id, na.rm = TRUE),
    n_distinct(crm$email_key, na.rm = TRUE),
    sum(!is.na(crm$company_name)),
    sum(!is.na(crm$membership_status)),
    sum(!is.na(crm$membership_plan))
  )
)

knitr::kable(crm_record_summary, align = c("l", "r"))
```

The table is based on `id`, `email_key`, `company_name`, `membership_status`
and `membership_plan`. It summarises the number of CRM contact
records and the availability of several key fields:

- **Contact records:** The CRM extract contains 6,051 contact records in total.
- **Distinct CRM IDs:** All 6,051 records have a unique CRM ID; no CRM ID is
  shared by multiple records.
- **Distinct normalised emails:** All 6,051 records have a unique email after
  converting emails to lowercase and removing unnecessary spaces.
- **Records with a school or organisation:** 3,934 contact records contain a
  non-missing value in `company_name`. This does not mean there are 3,934
  different schools or organisations, because multiple contacts can belong to
  the same organisation.
- **Records with membership status:** 1,038 contact records contain a value in
  `membership_status`. This is the total across all recorded status categories,
  not the number of Active members.
- **Records with membership plan:** 1,196 contact records contain a value in
  `membership_plan`. This counts contacts with a populated plan and does not
  represent the number of distinct membership-plan types.

These counts measure whether the relevant field is populated. They do not
establish that the recorded value is correct or current.

The record count is the number of rows in the CRM extract. It should not be
interpreted as the number of unique people until duplicate candidates have been
reviewed.

# 3 Duplicate analysis

## 3.1 Duplicate indicators

Email and CRM ID are treated as stronger identifiers. Matching first and last
names are treated only as manual-review candidates because different people can
share a name and one teacher may work across multiple schools.

```{r}
duplicate_summary <- tibble(
  duplicate_check = c(
    "Exact duplicate source rows",
    "Records with a duplicated CRM ID",
    "Records with a duplicated normalised email",
    "Contacts sharing a normalised first and last name"
  ),
  flagged_records = c(
    exact_duplicate_source_rows,
    sum(crm$duplicate_crm_id_candidate == "TRUE", na.rm = TRUE),
    sum(crm$duplicate_email_candidate == "TRUE", na.rm = TRUE),
    sum(crm$duplicate_name_candidate == "TRUE", na.rm = TRUE)
  )
) |>
  mutate(
    percentage_of_contacts = flagged_records / nrow(crm)
  )

duplicate_summary |>
  mutate(
    percentage_of_contacts = scales::percent(
      percentage_of_contacts,
      accuracy = 0.1
    )
  ) |>
  knitr::kable(align = c("l", "r", "r"))
```

The table is based on all 71 raw fields, `id`, `email_key` and `name_key`.
Exact rows compare all 71 original source fields.
CRM ID and email checks flag every record sharing the same identifier after
standardisation. The same-name result only flags contacts sharing a normalised
first and last name; these are review candidates, not confirmed duplicates.

## 3.2 Same-name review groups

```{r}
duplicate_name_groups <- crm |>
  filter(duplicate_name_candidate == "TRUE", !is.na(name_key)) |>
  group_by(name_key) |>
  summarise(
    contact_records = n(),
    distinct_emails = n_distinct(email_key, na.rm = TRUE),
    distinct_schools = n_distinct(school_name_key, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(contact_records), desc(distinct_schools), name_key)

duplicate_name_group_summary <- tibble(
  metric = c(
    "Duplicated-name groups",
    "Groups containing more than one email",
    "Groups associated with more than one school"
  ),
  groups = c(
    nrow(duplicate_name_groups),
    sum(duplicate_name_groups$distinct_emails > 1),
    sum(duplicate_name_groups$distinct_schools > 1)
  )
)

knitr::kable(duplicate_name_group_summary, align = c("l", "r"))
```

The table is based on `name_key`, `email_key` and `school_name_key`. A group
contains two or more contacts with the same
normalised first and last name. Distinct-email and distinct-school counts help
separate likely different people from records that warrant closer review, but
neither measure confirms identity.

These groups are not confirmed duplicates. The restricted review file retains
names, emails and organisations so the records can be checked manually.

# 4 Validity and consistency checks

## 4.1 Main checks

```{r}
validity_summary <- tibble(
  quality_check = c(
    "Missing CRM ID",
    "Missing email",
    "Invalid non-missing email format",
    "Full name differs from first plus last name",
    "Invalid non-missing Australian postcode",
    "Parser or row-structure issue"
  ),
  flagged_records = c(
    sum(is.na(crm$id)),
    sum(is.na(crm$email_key)),
    sum(crm$email_format_valid == "FALSE", na.rm = TRUE),
    sum(crm$full_name_component_mismatch == "TRUE", na.rm = TRUE),
    sum(crm$australian_postcode_format_valid == "FALSE", na.rm = TRUE),
    sum(crm$structure_issue == "TRUE", na.rm = TRUE)
  )
) |>
  mutate(percentage_of_contacts = flagged_records / nrow(crm))

validity_summary |>
  mutate(
    percentage_of_contacts = scales::percent(
      percentage_of_contacts,
      accuracy = 0.1
    )
  ) |>
  knitr::kable(align = c("l", "r", "r"))
```

The table is based on `id`, `email_key`, `email_format_valid`,
`full_name_component_mismatch`, `australian_postcode_format_valid` and
`structure_issue`. Each result counts records failing one technical
rule. A zero means no issue was detected by that rule, not that the underlying
data has been independently verified. A record may appear in more than one
check, so the rows should not be added together as a unique-contact total.

Email validity is a syntax check only. It does not confirm that the mailbox
exists, belongs to the named contact or can receive messages.

## 4.2 Basic school-name consistency check

This check identifies differences that collapse after case, spacing and
punctuation are standardised. Fuzzy matching and business-approved school-name
mapping are handled in `03_crm_school_groups.qmd`.

```{r}
school_name_variants <- crm |>
  filter(!is.na(company_name), !is.na(school_name_key)) |>
  distinct(school_name_key, company_name) |>
  count(school_name_key, name = "raw_name_variants") |>
  filter(raw_name_variants > 1) |>
  arrange(desc(raw_name_variants), school_name_key)

school_name_consistency_summary <- tibble(
  metric = c(
    "Distinct supplied school or organisation names",
    "Distinct basic-standardised names",
    "Standardised names with multiple supplied variants"
  ),
  result = c(
    n_distinct(crm$company_name, na.rm = TRUE),
    n_distinct(crm$school_name_key, na.rm = TRUE),
    nrow(school_name_variants)
  )
)

knitr::kable(school_name_consistency_summary, align = c("l", "r"))
```

The table is based on `company_name` and `school_name_key`. Supplied names are the distinct raw `company_name`
values. Standardised names remove differences in case, spacing and punctuation.
Multiple supplied variants under one key indicate formatting inconsistency, not
necessarily separate schools or confirmed duplicates.

# 5 Completeness of the 25 selected variables

## 5.1 Completeness calculation

Completeness is calculated after the missing-value cleaning in Report 01.
Whitespace-only values and agreed `No` placeholders are therefore counted as
missing where appropriate.

```{r}
profile_field <- function(field_name) {
  values <- crm[[field_name]]

  tibble(
    field = field_name,
    records_populated = sum(!is.na(values)),
    missing_records = sum(is.na(values)),
    completion_rate = sum(!is.na(values)) / nrow(crm),
    distinct_non_missing = n_distinct(values, na.rm = TRUE)
  )
}

crm_selected_field_completeness <- map_dfr(
  selected_business_fields,
  profile_field
) |>
  left_join(variable_contract, by = c("field" = "source_field")) |>
  arrange(position) |>
  select(
    position, field, records_populated, missing_records,
    completion_rate, distinct_non_missing
  )

crm_selected_field_completeness |>
  mutate(
    completion_rate = scales::percent(completion_rate, accuracy = 0.1)
  ) |>
  select(-position) |>
  knitr::kable(
    col.names = c(
      "Field", "Populated", "Missing", "Completion rate", "Distinct values"
    ),
    align = c("l", "r", "r", "r", "r")
  )
```

The table is based on the 25 fields listed in
`crm_analysis_variable_contract.csv`. For each field, completion rate is
populated records divided by all
6,051 CRM contacts after agreed blank and placeholder cleaning. It measures the
presence of a value only; it does not measure accuracy, relevance or whether the
field should apply to every contact.

## 5.2 Completeness chart

```{r}
crm_selected_field_completeness |>
  mutate(
    field = forcats::fct_reorder(field, completion_rate)
  ) |>
  ggplot(aes(x = completion_rate, y = field)) +
  geom_col(fill = "#315D80", width = 0.72) +
  geom_text(
    aes(label = scales::percent(completion_rate, accuracy = 0.1)),
    hjust = -0.08,
    size = 3
  ) +
  scale_x_continuous(
    limits = c(0, 1.08),
    breaks = seq(0, 1, 0.2),
    labels = scales::percent_format()
  ) +
  labs(
    x = "Completion rate",
    y = NULL,
    title = "Completeness of the 25 selected CRM variables"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.major.y = element_blank(),
    plot.title = element_text(face = "bold")
  )
```

## 5.3 Completeness bands

The bands organise attention; they do not determine whether a field is
business-critical. A low-completion optional field may be acceptable, while a
small gap in a required identifier may still need immediate action.

```{r}
crm_completeness_bands <- crm_selected_field_completeness |>
  mutate(
    completeness_band = case_when(
      completion_rate >= 0.95 ~ "95% or higher",
      completion_rate >= 0.75 ~ "75% to 94.9%",
      completion_rate >= 0.50 ~ "50% to 74.9%",
      TRUE ~ "Below 50%"
    )
  ) |>
  count(completeness_band, name = "fields") |>
  mutate(
    completeness_band = factor(
      completeness_band,
      levels = c(
        "95% or higher", "75% to 94.9%",
        "50% to 74.9%", "Below 50%"
      )
    )
  ) |>
  arrange(completeness_band)

knitr::kable(crm_completeness_bands, align = c("l", "r"))
```

The table is based on the `completion_rate` calculated for each selected field.
Each of the 25 selected variables is assigned to one
completion-rate band. These bands help prioritise review but are not pass/fail
standards, because optional and conditional fields can be legitimately blank.

# 6 Findings

```{r}
lowest_completeness <- crm_selected_field_completeness |>
  slice_min(completion_rate, n = 5, with_ties = FALSE) |>
  mutate(completion_label = scales::percent(completion_rate, accuracy = 0.1))

highest_completeness <- crm_selected_field_completeness |>
  slice_max(completion_rate, n = 5, with_ties = FALSE) |>
  mutate(completion_label = scales::percent(completion_rate, accuracy = 0.1))
```

The results should be interpreted using both data quality and business purpose:

- CRM ID and email quality determine whether records can be matched reliably.
- Same-name groups require manual review and should not be automatically merged.
- Low-completion fields indicate where collection processes may need review.
- Membership status completeness is reported here, but category agreement with
  membership tags is evaluated separately in Report 04.
- School-name checks in this report cover only simple formatting differences.

### Five least complete selected fields

```{r}
lowest_completeness |>
  select(field, records_populated, missing_records, completion_label) |>
  knitr::kable(
    col.names = c("Field", "Populated", "Missing", "Completion rate"),
    align = c("l", "r", "r", "r")
  )
```

The table is based on the five smallest values of `completion_rate`. These are
the five selected variables with the
lowest populated proportions. They identify where values are least available;
they do not show that the missing values are errors or that all five fields
should become mandatory.

# 7 Limitations

- The analysis describes the CRM extract dated 29 July 2026.
- Email format validation does not test mailbox deliverability.
- Same names are candidates for review, not confirmed duplicate people.
- Completeness measures whether a value exists, not whether it is correct or
  current.
- Only the 25 approved business variables are included in completeness results.
- Membership activity and relationship activity are different concepts and are
  not inferred in this report.

# 8 Save outputs

Aggregate results contain no contact-level personal information. Record-level
review files remain in the restricted folder.

```{r}
output_dir <- here("data", "analysis", "CRM", "02_data_quality")
restricted_dir <- file.path(output_dir, "restricted")

dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(restricted_dir, recursive = TRUE, showWarnings = FALSE)

duplicate_name_review_internal <- crm |>
  filter(duplicate_name_candidate == "TRUE") |>
  arrange(name_key, company_name, email) |>
  select(
    source_record_number, id, full_name, first_name, last_name, email,
    company_name, name_key, email_key, school_name_key
  )

data_quality_review_internal <- crm |>
  filter(
    structure_issue == "TRUE" |
      duplicate_crm_id_candidate == "TRUE" |
      duplicate_email_candidate == "TRUE" |
      email_format_valid == "FALSE" |
      full_name_component_mismatch == "TRUE" |
      australian_postcode_format_valid == "FALSE"
  ) |>
  select(
    source_record_number, id, full_name, first_name, last_name, email,
    postal_zip, company_name, structure_issue,
    duplicate_crm_id_candidate, duplicate_email_candidate,
    email_format_valid, full_name_component_mismatch,
    australian_postcode_format_valid
  )

aggregate_outputs <- list(
  crm_record_summary = crm_record_summary,
  crm_duplicate_summary = duplicate_summary,
  crm_duplicate_name_group_summary = duplicate_name_group_summary,
  crm_validity_summary = validity_summary,
  crm_school_name_consistency_summary = school_name_consistency_summary,
  crm_selected_field_completeness = crm_selected_field_completeness,
  crm_completeness_bands = crm_completeness_bands
)

iwalk(
  aggregate_outputs,
  ~ write_csv(.x, file.path(output_dir, paste0(.y, ".csv")), na = "")
)

write_csv(
  duplicate_name_review_internal,
  file.path(restricted_dir, "crm_duplicate_name_review_internal.csv"),
  na = ""
)

write_csv(
  data_quality_review_internal,
  file.path(restricted_dir, "crm_data_quality_review_internal.csv"),
  na = ""
)
```

# 9 Handoff

Report 02 completes the record-volume, duplicate, technical-validity and
selected-variable completeness analysis. The next report will focus on school
names, Group Admin, Group Member and Neither classifications.
