---
title: "CRM membership analysis"
subtitle: "Membership status tag reconciliation and review"
author: "Jordan"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: false
    code-fold: true
    embed-resources: true
execute:
  warning: false
  message: false
---

# Purpose

This report compares two CRM representations of membership:

1. The value recorded in `membership_status`.
2. The Active, On-hold and Expired categories detected in `tags`.

It reports category counts, measures the differences between the two sources and
creates restricted record-level files for manual review. Active membership is a
subscription classification and must not be interpreted as active engagement
with STAV.

# 1 Setup and input contract

```{r}
library(tidyverse)
library(lubridate)
library(here)

options(scipen = 999)
here::i_am("reports/CRM/04_crm_membership_analysis.qmd")

crm_file <- here(
  "data", "clean", "CRM", "restricted",
  "crm_contacts_analysis_internal.csv"
)

stopifnot(file.exists(crm_file))

crm <- read_csv(
  crm_file,
  col_types = cols(.default = col_character()),
  show_col_types = FALSE,
  progress = FALSE
)

required_fields <- c(
  "source_record_number", "id", "full_name", "email", "company_name",
  "school_name_key",
  "tags", "membership_status", "membership_status_clean",
  "membership_plan", "membership_plan_clean",
  "subscription_start_date_clean", "subscription_end_date_clean",
  "membership_tag_active", "membership_tag_on_hold",
  "membership_tag_expired", "membership_tag_status",
  "membership_tag_category_count", "multiple_membership_tag_categories",
  "membership_end_before_start", "active_membership_with_past_end_date"
)

stopifnot(all(required_fields %in% names(crm)))
stopifnot(!anyDuplicated(crm$source_record_number))

membership_categories <- c("active", "on-hold", "expired")
```

# 2 Membership status distribution

## 2.1 All recorded status categories

```{r}
membership_status_summary <- crm |>
  mutate(
    membership_status_display = replace_na(
      membership_status_clean,
      "No status recorded"
    )
  ) |>
  count(membership_status_display, name = "contact_records", sort = TRUE) |>
  mutate(percentage_of_contacts = contact_records / nrow(crm))

membership_status_summary |>
  mutate(
    percentage_of_contacts = scales::percent(
      percentage_of_contacts,
      accuracy = 0.1
    )
  ) |>
  knitr::kable(
    col.names = c("Membership status", "Contacts", "Percentage"),
    align = c("l", "r", "r")
  )
```

The table is based on `membership_status_clean`. Each contact contributes to one cleaned
`membership_status` category or to `No status recorded`. The counts describe
the CRM field as stored; they do not confirm payment, current eligibility or
recent engagement.

## 2.2 Status counts used in the three-category comparison

```{r}
membership_status_three_category <- crm |>
  filter(membership_status_clean %in% membership_categories) |>
  count(
    category = membership_status_clean,
    name = "status_records"
  ) |>
  complete(
    category = membership_categories,
    fill = list(status_records = 0L)
  ) |>
  mutate(category = factor(category, levels = membership_categories)) |>
  arrange(category)

knitr::kable(
  membership_status_three_category,
  col.names = c("Category", "Records from membership status"),
  align = c("l", "r")
)
```

The table is based on `membership_status_clean` after filtering to Active,
On-hold and Expired. It keeps only these three categories so
they can be compared directly with the three requested tag categories.
Cancelled, Pending, Paused and Pending-cancel remain in the full status table
but are not forced into one of these categories.

# 3 Membership tag distribution

## 3.1 Priority classification requested for tags

Report 01 applies the agreed order:

```r
case_when(
  str_detect(tags, "\\[MEMBERSHIP\\] ACTIVE") ~ "active",
  str_detect(tags, "\\[MEMBERSHIP\\] ON-HOLD") ~ "on-hold",
  str_detect(tags, "\\[MEMBERSHIP\\] EXPIRED") ~ "expired"
)
```

If a contact contains more than one of these tags, Active takes priority over
On-hold, and On-hold takes priority over Expired.

```{r}
membership_tag_classification_summary <- crm |>
  mutate(
    membership_tag_display = replace_na(
      membership_tag_status,
      "No selected membership tag"
    )
  ) |>
  count(membership_tag_display, name = "contact_records", sort = TRUE) |>
  mutate(percentage_of_contacts = contact_records / nrow(crm))

membership_tag_classification_summary |>
  mutate(
    percentage_of_contacts = scales::percent(
      percentage_of_contacts,
      accuracy = 0.1
    )
  ) |>
  knitr::kable(
    col.names = c("Membership tag category", "Contacts", "Percentage"),
    align = c("l", "r", "r")
  )
```

The table is based on `membership_tag_status`, which is derived from `tags`.
Each contact receives at most one tag category using
the agreed priority order. When several selected tags occur together, Active
overrides On-hold and Expired, and On-hold overrides Expired. The table is
therefore mutually exclusive but can hide lower-priority tag occurrences.

## 3.2 Independent tag occurrence counts

Independent counts do not apply category priority. A contact carrying more than
one selected membership tag contributes to each matching tag count.

```{r}
membership_tag_occurrence_summary <- tibble(
  category = membership_categories,
  tag_records = c(
    sum(crm$membership_tag_active == "TRUE", na.rm = TRUE),
    sum(crm$membership_tag_on_hold == "TRUE", na.rm = TRUE),
    sum(crm$membership_tag_expired == "TRUE", na.rm = TRUE)
  )
)

knitr::kable(
  membership_tag_occurrence_summary,
  col.names = c("Tag category", "Contacts carrying tag"),
  align = c("l", "r")
)
```

The table is based on `membership_tag_active`, `membership_tag_on_hold` and
`membership_tag_expired`. Each selected tag is counted independently. A
contact carrying two or three selected membership tags contributes to every
matching row, so these values can exceed the mutually exclusive classification
counts above.

## 3.3 Multiple selected membership tags

```{r}
multiple_membership_tag_summary <- crm |>
  filter(multiple_membership_tag_categories == "TRUE") |>
  count(
    membership_tag_active,
    membership_tag_on_hold,
    membership_tag_expired,
    name = "contact_records",
    sort = TRUE
  )

tibble(
  metric = "Contacts carrying more than one selected membership tag",
  contact_records = sum(
    crm$multiple_membership_tag_categories == "TRUE",
    na.rm = TRUE
  )
) |>
  knitr::kable(align = c("l", "r"))

knitr::kable(
  multiple_membership_tag_summary,
  col.names = c("Active tag", "On-hold tag", "Expired tag", "Contacts"),
  align = c("c", "c", "c", "r")
)
```

The tables are based on `membership_tag_category_count` and the three tag flags.
The first table gives the unique number of
contacts carrying more than one selected membership tag. The second shows which
tag combinations occur. Multiple tags may reflect history or migration and
require review before the priority category is treated as authoritative.

# 4 Compare membership status with tag classification

## 4.1 Category-count comparison

```{r}
membership_tag_three_category <- crm |>
  filter(membership_tag_status %in% membership_categories) |>
  count(
    category = membership_tag_status,
    name = "tag_classification_records"
  ) |>
  complete(
    category = membership_categories,
    fill = list(tag_classification_records = 0L)
  ) |>
  mutate(category = factor(category, levels = membership_categories)) |>
  arrange(category)

membership_count_comparison <- membership_status_three_category |>
  left_join(membership_tag_three_category, by = "category") |>
  mutate(
    difference_tag_minus_status =
      tag_classification_records - status_records,
    absolute_difference = abs(difference_tag_minus_status)
  )

knitr::kable(
  membership_count_comparison,
  col.names = c(
    "Category", "Membership status", "Tag classification",
    "Tag minus status", "Absolute difference"
  ),
  align = c("l", "r", "r", "r", "r")
)
```

The table is based on `membership_status_clean` and `membership_tag_status`.
For each of the three categories, `Tag minus status` compares category totals; a
negative result means fewer contacts in the priority tag classification than in
the status field. It does not identify the affected contacts or distinguish
missing tags from conflicting categories.

The difference in category totals does not identify which records disagree.
The record-level comparison below separates missing classifications from true
category conflicts.

## 4.2 Record-level comparison classes

```{r}
crm_membership_comparison <- crm |>
  mutate(
    status_in_comparison = membership_status_clean %in% membership_categories,
    tag_in_comparison = membership_tag_status %in% membership_categories,
    membership_comparison_class = case_when(
      status_in_comparison & tag_in_comparison &
        membership_status_clean == membership_tag_status ~
        "Status and tag agree",
      status_in_comparison & tag_in_comparison &
        membership_status_clean != membership_tag_status ~
        "Status and tag disagree",
      status_in_comparison & !tag_in_comparison ~
        "Selected status but no selected tag",
      !status_in_comparison & tag_in_comparison ~
        "Selected tag but no comparable status",
      TRUE ~ "Neither source has a comparable category"
    )
  )

membership_comparison_summary <- crm_membership_comparison |>
  count(membership_comparison_class, name = "contact_records", sort = TRUE) |>
  mutate(percentage_of_contacts = contact_records / nrow(crm_membership_comparison))

membership_comparison_summary |>
  mutate(
    percentage_of_contacts = scales::percent(
      percentage_of_contacts,
      accuracy = 0.1
    )
  ) |>
  knitr::kable(
    col.names = c("Comparison result", "Contacts", "Percentage"),
    align = c("l", "r", "r")
  )
```

The table is based on a row-by-row comparison of `membership_status_clean` and
`membership_tag_status`. Agreement requires the cleaned status and priority
tag classification to be the same selected category. Missing counterparts and
true category disagreements are shown separately. Categories outside Active,
On-hold and Expired are treated as not directly comparable.

The calculation starts with one contact at a time and asks two questions:

1. Is `membership_status_clean` Active, On-hold or Expired?
2. Is `membership_tag_status` Active, On-hold or Expired?

The contact is then assigned to exactly one result:

- If both answers are yes and the values are identical, the result is `Status
  and tag agree`.
- If both answers are yes but the values differ, the result is `Status and tag
  disagree`.
- If only the status is one of the three categories, the result is `Selected
  status but no selected tag`.
- If only the tag is one of the three categories, the result is `Selected tag
  but no comparable status`.
- If neither field contains one of the three comparable categories, the result
  is `Neither source has a comparable category`.

For example, Active status plus Active tag is an agreement; Active status plus
Expired tag is a disagreement; Active status with no selected membership tag
is a missing-tag case. A Cancelled status with no selected membership tag is
placed in the final class because Cancelled has no equivalent among the three
selected tag categories.

## 4.3 Status by tag cross-table

```{r}
membership_status_tag_crosstab <- crm_membership_comparison |>
  mutate(
    status_display = replace_na(
      membership_status_clean,
      "No status recorded"
    ),
    tag_display = replace_na(
      membership_tag_status,
      "No selected membership tag"
    )
  ) |>
  count(status_display, tag_display, name = "contact_records") |>
  arrange(desc(contact_records), status_display, tag_display)

membership_status_tag_matrix <- membership_status_tag_crosstab |>
  pivot_wider(
    names_from = tag_display,
    values_from = contact_records,
    values_fill = 0
  )

knitr::kable(membership_status_tag_matrix, align = "r")
```

The table is based on a cross-tabulation of `membership_status_clean` and
`membership_tag_status`. Rows are cleaned membership statuses and columns are
priority tag classifications. Cells on matching Active, On-hold and Expired
positions represent agreement; off-diagonal cells show differences. Missing
categories expose records represented by only one source.

# 5 Membership plan, organisation and date checks

## 5.1 Membership group

```{r}
school_membership_plans <- c(
  "Secondary School - Group Membership",
  "P - 12 School - Group Membership",
  "Primary School - Group Membership"
)

organisation_membership_plans <- "Organisation - Group Membership"

crm_membership_group <- crm |>
  mutate(
    membership_group = case_when(
      membership_plan_clean %in% organisation_membership_plans ~
        "Organisation",
      membership_plan_clean %in% school_membership_plans ~
        "School membership",
      !is.na(membership_plan_clean) ~ "Individual",
      TRUE ~ "Blank"
    ),
    membership_group = factor(
      membership_group,
      levels = c(
        "Organisation", "School membership", "Individual", "Blank"
      )
    )
  )

membership_group_summary <- crm_membership_group |>
  count(membership_group, name = "contact_records", .drop = FALSE) |>
  mutate(percentage_of_contacts = contact_records / nrow(crm_membership_group))

membership_group_summary |>
  mutate(
    percentage_of_contacts = scales::percent(
      percentage_of_contacts,
      accuracy = 0.1
    )
  ) |>
  knitr::kable(
    col.names = c("Membership group", "Contacts", "Percentage"),
    align = c("l", "r", "r")
  )
```

The table is based only on `membership_plan_clean`, so each contact is counted
once. `Organisation` means the plan is Organisation - Group Membership.
`School membership` combines Primary School, Secondary School and P-12 School
Group Membership plans. `Individual` includes every other non-blank plan, such
as teacher, student teacher, academic, technician, life and retired plans.
`Blank` means no membership plan is recorded. Membership status is not used in
this classification, so membership type and membership activity are not mixed.

```{r}
membership_plan_distinct_summary <- crm |>
  mutate(
    membership_plan_display = if_else(
      is.na(membership_plan),
      "Blank",
      membership_plan
    )
  ) |>
  count(membership_plan_display, name = "contact_records", sort = TRUE) |>
  mutate(percentage_of_contacts = contact_records / nrow(crm))

membership_plan_distinct_summary |>
  mutate(
    percentage_of_contacts = scales::percent(
      percentage_of_contacts,
      accuracy = 0.1
    )
  ) |>
  knitr::kable(
    col.names = c("Distinct membership plan", "Contacts", "Percentage"),
    align = c("l", "r", "r")
  )
```

This second table is based directly on the original `membership_plan` field,
not only `membership_plan_clean`. Each distinct supplied plan value is shown
separately, and missing values are displayed as Blank. It provides the detailed
plan counts underlying the four broader membership groups above.

## 5.2 Organisation-level membership table

```{r}
organisation_membership <- crm_membership_group |>
  filter(!is.na(school_name_key)) |>
  group_by(school_name_key) |>
  summarise(
    organization_name = {
      name_counts <- tibble(name = company_name) |>
        filter(!is.na(name)) |>
        count(name, sort = TRUE)
      if (nrow(name_counts) == 0) school_name_key else name_counts$name[[1]]
    },
    organization_status = if_else(
      any(membership_status_clean == "active", na.rm = TRUE),
      "Active",
      "Not active"
    ),
    organization_plan = {
      group_plans <- sort(unique(membership_plan_clean[
        membership_plan_clean %in% c(
          organisation_membership_plans,
          school_membership_plans
        )
      ]))
      if (length(group_plans) == 0) {
        "No group membership plan"
      } else {
        paste(group_plans, collapse = "; ")
      }
    },
    organization_type = case_when(
      any(membership_plan_clean %in% school_membership_plans) ~ "School",
      any(membership_plan_clean %in% organisation_membership_plans) ~
        "Organisation",
      TRUE ~ "Unknown"
    ),
    .groups = "drop"
  ) |>
  select(
    `Organization name` = organization_name,
    `Organization status` = organization_status,
    `Organization plan` = organization_plan,
    `Organization type` = organization_type
  ) |>
  arrange(`Organization name`)

organisation_membership_total <- tibble(
  metric = "Total organisations",
  organisations = nrow(organisation_membership)
)

organisation_status_summary <- organisation_membership |>
  count(`Organization status`, name = "organisations")

organisation_type_summary <- organisation_membership |>
  count(`Organization type`, name = "organisations")

organisation_membership_total |>
  knitr::kable(
    col.names = c("Metric", "Organisations"),
    align = c("l", "r")
  )

organisation_status_summary |>
  knitr::kable(
    col.names = c("Organization status", "Organisations"),
    align = c("l", "r")
  )

organisation_type_summary |>
  knitr::kable(
    col.names = c("Organization type", "Organisations"),
    align = c("l", "r")
  )

organisation_membership |>
  slice_head(n = 10) |>
  knitr::kable(align = c("l", "l", "l", "l"))
```

The summary tables count all organisations, first as a total and then by
`Organization status` and `Organization type`. These two breakdowns describe
different attributes and should not be added together. The detail table displays
the first 10 organisations in alphabetical order as a preview; the complete
organisation-level dataset is saved as a CSV file. The complete table has one
row per non-blank `school_name_key`. `Organization name` is
the most frequently occurring non-blank `company_name` for that standardised
key. `Organization status` is Active when at least one associated contact has
`membership_status_clean` equal to Active; otherwise it is Not active,
including organisations with no recorded membership. `Organization plan`
lists the distinct school or organisation group plans recorded among associated
contacts; if none is found, it is shown as No group membership plan.
`Organization type` is School when any school group plan is found,
Organisation when an Organisation - Group Membership plan is found and no
school group plan is present, and Unknown when neither plan type is available.

## 5.3 Membership date and plan coverage

These checks support membership reconciliation. They do not establish payment
status or relationship activity.

```{r}
membership_supporting_checks <- tibble(
  check = c(
    "Membership status recorded",
    "Membership plan recorded",
    "Subscription start date recorded",
    "Subscription end date recorded",
    "Membership end date before start date",
    "Active status with past end date"
  ),
  contact_records = c(
    sum(!is.na(crm$membership_status_clean)),
    sum(!is.na(crm$membership_plan_clean)),
    sum(!is.na(crm$subscription_start_date_clean)),
    sum(!is.na(crm$subscription_end_date_clean)),
    sum(crm$membership_end_before_start == "TRUE", na.rm = TRUE),
    sum(crm$active_membership_with_past_end_date == "TRUE", na.rm = TRUE)
  )
)

knitr::kable(membership_supporting_checks, align = c("l", "r"))
```

The table is based on `membership_status_clean`, `membership_plan_clean`, the
clean subscription dates and the two date-validation flags. The first four rows measure whether supporting
membership values are populated. The final two rows apply date-consistency
rules. An Active status with a past end date is a review flag, not automatic
proof that the membership is inactive.

# 6 Findings and interpretation

- `membership_status` is the primary status field, while tags provide a second
  classification that may lag or reflect migration history.
- Matching totals do not prove record-level agreement; the cross-table and
  review file identify the actual differences.
- Contacts with multiple selected membership tags require review before the
  priority classification is treated as definitive.
- Cancelled, pending, paused and pending-cancel statuses have no direct category
  in the requested three-tag comparison.
- Active membership means a membership record is labelled active. It does not
  measure whether a person has recently engaged with STAV.

# 7 Save outputs

```{r}
output_dir <- here("data", "analysis", "CRM", "04_membership")
restricted_dir <- file.path(output_dir, "restricted")

dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(restricted_dir, recursive = TRUE, showWarnings = FALSE)

membership_difference_review_internal <- crm_membership_comparison |>
  filter(membership_comparison_class != "Status and tag agree") |>
  filter(
    membership_comparison_class !=
      "Neither source has a comparable category" |
      !is.na(membership_status_clean) |
      !is.na(membership_plan_clean) |
      multiple_membership_tag_categories == "TRUE"
  ) |>
  select(
    source_record_number, id, full_name, email, company_name,
    membership_status, membership_status_clean,
    membership_plan, membership_plan_clean, tags,
    membership_tag_active, membership_tag_on_hold,
    membership_tag_expired, membership_tag_status,
    multiple_membership_tag_categories,
    membership_comparison_class,
    subscription_start_date_clean, subscription_end_date_clean,
    active_membership_with_past_end_date
  ) |>
  arrange(membership_comparison_class, membership_status_clean, full_name)

multiple_membership_tags_review_internal <- crm_membership_comparison |>
  filter(multiple_membership_tag_categories == "TRUE") |>
  select(
    source_record_number, id, full_name, email, company_name,
    membership_status, membership_status_clean, tags,
    membership_tag_active, membership_tag_on_hold,
    membership_tag_expired, membership_tag_status,
    membership_comparison_class
  ) |>
  arrange(full_name)

aggregate_outputs <- list(
  crm_membership_status_summary = membership_status_summary,
  crm_membership_status_three_category = membership_status_three_category,
  crm_membership_tag_classification_summary =
    membership_tag_classification_summary,
  crm_membership_tag_occurrence_summary = membership_tag_occurrence_summary,
  crm_multiple_membership_tag_summary = multiple_membership_tag_summary,
  crm_membership_count_comparison = membership_count_comparison,
  crm_membership_comparison_summary = membership_comparison_summary,
  crm_membership_status_tag_crosstab = membership_status_tag_crosstab,
  crm_membership_status_tag_matrix = membership_status_tag_matrix,
  crm_membership_supporting_checks = membership_supporting_checks,
  crm_membership_group_summary = membership_group_summary,
  crm_membership_plan_distinct_summary = membership_plan_distinct_summary,
  crm_organisation_membership_total = organisation_membership_total,
  crm_organisation_status_summary = organisation_status_summary,
  crm_organisation_type_summary = organisation_type_summary,
  crm_organisation_membership = organisation_membership
)

iwalk(
  aggregate_outputs,
  ~ write_csv(.x, file.path(output_dir, paste0(.y, ".csv")), na = "")
)

write_csv(
  membership_difference_review_internal,
  file.path(restricted_dir, "crm_membership_difference_review_internal.csv"),
  na = ""
)

write_csv(
  multiple_membership_tags_review_internal,
  file.path(restricted_dir, "crm_multiple_membership_tags_review_internal.csv"),
  na = ""
)

write_csv(
  crm_membership_group |>
    mutate(membership_group = as.character(membership_group)) |>
    select(source_record_number, id, membership_plan_clean, membership_group),
  file.path(restricted_dir, "crm_contact_membership_group_internal.csv"),
  na = ""
)
```

# 8 Handoff

This report completes the requested comparison between `membership_status` and
the three selected membership tag categories. Report 05 will address active and
dormant relationships only after the engagement definition and required
activity data are confirmed.
