---
title: "CRM health check summary and recommendations"
subtitle: "Data quality school groups and membership reconciliation"
author: "Jordan"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    number-sections: false
    code-fold: true
    embed-resources: true
execute:
  warning: false
  message: false
---

# Executive summary

This report consolidates the completed CRM health-check work from the data
quality, school-group and membership reports. The analysis uses the 29 July
2026 CRM extract and the 25 agreed business variables.

The CRM contains 6,051 contact records with unique CRM IDs and unique normalised
email addresses. No exact duplicate rows, duplicate CRM IDs or duplicate emails
were found. However, 1,379 contacts belong to same-name groups and require
manual review rather than automatic merging.

The main issues are incomplete optional and segmentation fields, inconsistent
school or organisation names, group-role contacts without an organisation, and
differences between membership status and membership tags. Active and dormant
relationship analysis is not included because the selected variables do not
contain a reliable last-contact or engagement measure.

# 1 Scope and source reports

```{r}
library(tidyverse)
library(here)

options(scipen = 999)
here::i_am("reports/CRM/06_crm_summary_recommendations.qmd")

quality_dir <- here("data", "analysis", "CRM", "02_data_quality")
groups_dir <- here("data", "analysis", "CRM", "03_school_groups")
membership_dir <- here("data", "analysis", "CRM", "04_membership")

required_files <- c(
  file.path(quality_dir, "crm_record_summary.csv"),
  file.path(quality_dir, "crm_duplicate_summary.csv"),
  file.path(quality_dir, "crm_duplicate_name_group_summary.csv"),
  file.path(quality_dir, "crm_validity_summary.csv"),
  file.path(quality_dir, "crm_selected_field_completeness.csv"),
  file.path(groups_dir, "crm_group_tag_presence_summary.csv"),
  file.path(groups_dir, "crm_school_group_coverage.csv"),
  file.path(groups_dir, "crm_missing_school_group_summary.csv"),
  file.path(groups_dir, "crm_school_name_variant_summary.csv"),
  file.path(groups_dir, "crm_school_name_fuzzy_candidates.csv"),
  file.path(membership_dir, "crm_membership_status_summary.csv"),
  file.path(membership_dir, "crm_membership_count_comparison.csv"),
  file.path(membership_dir, "crm_membership_comparison_summary.csv"),
  file.path(membership_dir, "crm_membership_supporting_checks.csv"),
  file.path(membership_dir, "crm_membership_group_summary.csv"),
  file.path(membership_dir, "crm_membership_plan_distinct_summary.csv"),
  file.path(membership_dir, "crm_organisation_membership_total.csv"),
  file.path(membership_dir, "crm_organisation_status_summary.csv"),
  file.path(membership_dir, "crm_organisation_type_summary.csv"),
  file.path(membership_dir, "crm_organisation_membership.csv"),
  file.path(groups_dir, "crm_school_name_standardisation_summary.csv"),
  file.path(groups_dir, "crm_school_name_standardisation_examples.csv"),
  file.path(groups_dir, "crm_school_name_review_method_summary.csv")
)

stopifnot(all(file.exists(required_files)))

record_summary <- read_csv(required_files[1], show_col_types = FALSE)
duplicate_summary <- read_csv(required_files[2], show_col_types = FALSE)
duplicate_name_summary <- read_csv(required_files[3], show_col_types = FALSE)
validity_summary <- read_csv(required_files[4], show_col_types = FALSE)
field_completeness <- read_csv(required_files[5], show_col_types = FALSE)
group_tag_presence_summary <- read_csv(required_files[6], show_col_types = FALSE)
school_group_coverage <- read_csv(required_files[7], show_col_types = FALSE)
missing_school_groups <- read_csv(required_files[8], show_col_types = FALSE)
school_variant_summary <- read_csv(required_files[9], show_col_types = FALSE)
school_fuzzy_candidates <- read_csv(required_files[10], show_col_types = FALSE)
membership_status_summary <- read_csv(required_files[11], show_col_types = FALSE)
membership_count_comparison <- read_csv(required_files[12], show_col_types = FALSE)
membership_comparison_summary <- read_csv(required_files[13], show_col_types = FALSE)
membership_supporting_checks <- read_csv(required_files[14], show_col_types = FALSE)
membership_group_summary <- read_csv(required_files[15], show_col_types = FALSE)
membership_plan_distinct_summary <- read_csv(required_files[16], show_col_types = FALSE)
organisation_membership_total <- read_csv(required_files[17], show_col_types = FALSE)
organisation_status_summary <- read_csv(required_files[18], show_col_types = FALSE)
organisation_type_summary <- read_csv(required_files[19], show_col_types = FALSE)
organisation_membership <- read_csv(required_files[20], show_col_types = FALSE)
school_name_standardisation_summary <- read_csv(required_files[21], show_col_types = FALSE)
school_name_standardisation_examples <- read_csv(required_files[22], show_col_types = FALSE)
school_name_review_method_summary <- read_csv(required_files[23], show_col_types = FALSE)
```

Completed reports:

- Report 01 prepares and validates the reusable CRM data.
- Report 02 analyses record volume, duplicates, validity and completeness.
- Report 03 analyses school names and Group Admin or Group Member coverage.
- Report 04 compares membership status with membership tags.
- Report 05 is intentionally omitted because no reliable contact-activity data
  is available.

# 2 Status of the CRM health-check questions

```{r}
crm_health_check_status <- tribble(
  ~question, ~status, ~current_result,
  "How many records do we have?", "Completed",
  "6,051 CRM contact records.",
  "How many are duplicates?", "Completed",
  "No exact rows, CRM IDs or emails are duplicated; 1,379 same-name contacts require review.",
  "How complete is the data?", "Completed",
  "Completeness is reported for all 25 approved variables.",
  "How accurate or current is the data?", "Partially completed",
  "Technical checks are complete; real-world accuracy and recency require business review.",
  "How many people are genuinely active?", "Not measurable",
  "The selected data has no reliable last-contact or engagement field.",
  "How many people are dormant?", "Not measurable",
  "The selected data has no reliable last-contact or engagement field.",
  "Can audiences be segmented reliably?", "Partially completed",
  "Available segmentation fields are profiled, but several have low completeness.",
  "Can engagement be measured per record?", "Not measurable",
  "No dated event, email or website engagement data is included."
)

knitr::kable(
  crm_health_check_status,
  col.names = c("Health-check question", "Status", "Current result"),
  align = c("l", "l", "l")
)
```

The table is based on the outputs of Reports 02, 03 and 04. `Completed` means the available 25-variable extract
supports the requested calculation. `Partially completed` means technical
evidence is available but business confirmation is still required. `Not
measurable` means the necessary activity field is absent and no proxy has been
substituted.

# 3 Data quality findings

## 3.1 Record and duplicate summary

```{r}
record_summary |>
  knitr::kable(
    col.names = c("Metric", "Records"),
    align = c("l", "r")
  )

duplicate_summary |>
  mutate(
    percentage_of_contacts = scales::percent(
      percentage_of_contacts,
      accuracy = 0.1
    )
  ) |>
  knitr::kable(
    col.names = c("Duplicate check", "Flagged records", "Percentage"),
    align = c("l", "r", "r")
  )
```

The first table is based on `id`, `email_key`, `company_name`,
`membership_status` and `membership_plan`. Each row uses the following logic:

- **Contact records:** The CRM extract contains 6,051 contact records in total.
- **Distinct CRM IDs:** All 6,051 records have a unique CRM ID.
- **Distinct normalised emails:** All 6,051 records have a unique email after
  lowercase and whitespace standardisation.
- **Records with a school or organisation:** 3,934 contact records contain a
  non-missing `company_name`. This is a contact count, not a count of unique
  organisations.
- **Records with membership status:** 1,038 contact records contain a value in
  `membership_status`. This includes all status categories and is not the number
  of Active members.
- **Records with membership plan:** 1,196 contact records contain a value in
  `membership_plan`. This is not the number of distinct plan types.

The second table applies four duplicate checks. Same-name records are review
candidates only; the zero ID and email results do not prove that the same person
never used two different identifiers.

No records should be merged solely because first and last names match. The 566
same-name groups all contain more than one email, and 155 groups are associated
with more than one school or organisation.

## 3.2 Technical validity

```{r}
validity_summary |>
  mutate(
    percentage_of_contacts = scales::percent(
      percentage_of_contacts,
      accuracy = 0.1
    )
  ) |>
  knitr::kable(
    col.names = c("Quality check", "Flagged records", "Percentage"),
    align = c("l", "r", "r")
  )
```

The table is based on the validation flags created in Report 01. Each row counts records failing one technical rule.
A zero indicates no problem detected by that rule, not independent verification
of the real-world value. Rows may overlap and should not be summed as unique
contacts.

CRM IDs and email formats are complete under the available technical checks.
The immediate technical review population consists of four invalid non-missing
postcodes and three parser-affected records.

## 3.3 Completeness

```{r}
field_completeness |>
  mutate(
    completion_rate = scales::percent(completion_rate, accuracy = 0.1)
  ) |>
  arrange(position) |>
  select(field, records_populated, missing_records, completion_rate) |>
  knitr::kable(
    col.names = c("Field", "Populated", "Missing", "Completion rate"),
    align = c("l", "r", "r", "r")
  )
```

The table is based on the 25 approved source fields. For each field, completion rate is the share of all 6,051 contacts
with a populated value after cleaning. It measures presence only, not accuracy
or whether a field should apply to every type of contact.

Thirteen of the 25 approved variables are below 50% complete. Low completion
does not automatically mean poor quality: membership fields may be blank for
non-members, and the leadership `Other` field should be blank unless `Other` is
selected. STAV should identify which fields are required for each contact type
before setting completeness targets.

# 4 School and group findings

## 4.1 Contact group roles

```{r}
group_tag_presence_summary |>
  mutate(
    percentage_of_contacts = scales::percent(
      percentage_of_contacts,
      accuracy = 0.1
    )
  ) |>
  knitr::kable(
    col.names = c("Tag presence", "Contacts", "Percentage"),
    align = c("l", "r", "r")
  )
```

This table is based on the Group Admin and Group Member flags derived from
`tags`. The Group Admin count includes Admin-only and Both contacts; the Group
Member count includes Member-only and Both contacts. The totals overlap and
must not be added together as a unique-contact total. Each percentage uses all
6,051 CRM contacts as the denominator, and the overlapping rows therefore do
not sum to 100%.

## 4.2 Organisation-level coverage

```{r}
school_group_coverage |>
  knitr::kable(
    col.names = c(
      "Population", "Distinct organisations",
      "Organisations with no Group Admin",
      "Organisations with multiple Group Admins",
      "Organisations with no Group Member",
      "Organisations containing Neither contacts only"
    ),
    align = c("l", "r", "r", "r", "r", "r")
  )
```

The tables are based on `company_name` and the Group role flags derived from
`tags`. Coverage indicators identify organisations that
may require role review; they are not automatic errors. The second table counts
group-role contacts with no organisation, who cannot yet be included reliably
in an organisation-level assessment.

Among 323 organisations with at least one group-role tag, 46 have no Group
Admin, 15 have multiple Group Admins and 136 have no Group Member. In addition,
183 contacts carry a group-role tag but have no `company_name`, so they cannot
be allocated reliably to an organisation.

## 4.3 School-name consistency

```{r}
school_name_standardisation_examples |>
  knitr::kable(align = c("l", "l", "r"))
```

The table shows 10 actual transformations from the 1,487 distinct supplied
`company_name` values to the standardised `school_name_key` values. Across the
full data, basic cleaning reduces 1,487 supplied names to 1,301 distinct keys, a
reduction of 186. The cleaning function converts text to
lowercase, removes leading, trailing and repeated whitespace, removes
characters other than letters, numbers and spaces, and collapses whitespace
again. It does not expand abbreviations, correct spelling or remove words such
as School, College or Campus.

```{r}
school_name_review_method_summary |>
  knitr::kable(
    col.names = c("Review measure", "Result", "Calculation rule"),
    align = c("l", "r", "l")
  )
```

The 156 variant keys are obtained by grouping records by `school_name_key`,
counting distinct supplied `company_name` values within each key and retaining
keys with more than one supplied value. The 489 fuzzy candidates are pairs of
different keys, not 489 confirmed duplicates. Keys shorter than six characters
are excluded, remaining names must share the same first three characters, and
a pair is retained when its normalised Levenshtein edit distance is no more
than 20% or one standardised name contains the other. These candidates require
manual review because similar names may represent different campuses or
organisations.


not finish

# 5 Membership findings

## 5.1 Membership status and tag totals

```{r}
membership_count_comparison |>
  knitr::kable(
    col.names = c(
      "Category", "Membership status", "Tag classification",
      "Tag minus status", "Absolute difference"
    ),
    align = c("l", "r", "r", "r", "r")
  )
```

The table is based on `membership_status_clean` and `membership_tag_status`.
Status and tag counts are independently calculated
for the same three categories. `Tag minus status` shows the net count
difference, but record-level review is required to distinguish missing values
from category conflicts.

The tag classification follows the requested Active, On-hold, Expired priority
order. Independent tag counts are available in Report 04 for contacts carrying
multiple selected membership tags.

## 5.2 Record-level agreement

```{r}
membership_comparison_summary |>
  mutate(
    percentage_of_contacts = scales::percent(
      percentage_of_contacts,
      accuracy = 0.1
    )
  ) |>
  knitr::kable(
    col.names = c("Comparison result", "Contacts", "Percentage"),
    align = c("l", "r", "r")
  )

membership_supporting_checks |>
  knitr::kable(
    col.names = c("Supporting check", "Contacts"),
    align = c("l", "r")
  )
```

The first table compares `membership_status_clean` with
`membership_tag_status` separately for every contact. Only Active, On-hold and
Expired are treated as comparable because they are the three categories that
exist in both the selected status and tag analysis.

For each contact, the calculation first asks whether the status is one of these
three categories and whether the tag classification is one of these three
categories. The contact is then placed into exactly one result:

- `Status and tag agree`: both fields contain a comparable category and the two
  values are the same, such as Active plus Active.
- `Status and tag disagree`: both fields contain a comparable category but the
  values differ, such as Active plus Expired.
- `Selected status but no selected tag`: the status is Active, On-hold or
  Expired, but no comparable tag classification exists.
- `Selected tag but no comparable status`: a selected membership tag exists,
  but status is blank or contains another value such as Cancelled.
- `Neither source has a comparable category`: neither field contains Active,
  On-hold or Expired. This does not necessarily mean both original fields are
  blank; for example, Cancelled status with no selected tag belongs here.

The five rows are mutually exclusive and add to all 6,051 contacts. The second
table is separate: it reports whether plan and date fields are populated and
flags date inconsistencies. Active with a past end date is a review condition,
not an automatic membership decision.

The main review populations are:

- 90 contacts with an Active, On-hold or Expired status but no corresponding
  selected tag.
- 32 contacts with a selected tag but no directly comparable status.
- 11 contacts whose selected status and priority tag classification disagree.
- 11 contacts carrying more than one selected membership tag.
- 49 contacts labelled Active whose subscription end date precedes the extract
  date.

## 5.3 Membership group

```{r}
membership_group_summary |>
  mutate(
    percentage_of_contacts = scales::percent(
      percentage_of_contacts,
      accuracy = 0.1
    )
  ) |>
  knitr::kable(
    col.names = c("Membership group", "Contacts", "Percentage"),
    align = c("l", "r", "r")
  )
```

The table is based only on `membership_plan_clean`, and every contact is counted
once. `Organisation` is Organisation - Group Membership. `School membership`
combines Primary School, Secondary School and P-12 School Group Membership.
`Individual` contains every other non-blank plan, including teacher, student
teacher, academic, technician, life and retired plans. `Blank` means no
membership plan is recorded. Membership status is deliberately not used, so
membership type and membership activity remain separate concepts.

```{r}
membership_plan_distinct_summary |>
  mutate(
    percentage_of_contacts = scales::percent(
      percentage_of_contacts,
      accuracy = 0.1
    )
  ) |>
  knitr::kable(
    col.names = c("Distinct membership plan", "Contacts", "Percentage"),
    align = c("l", "r", "r")
  )
```

This second table is based directly on the original `membership_plan` field,
not only `membership_plan_clean`. It shows every distinct supplied plan value
and displays missing values as Blank. These detailed plan counts are the source
values underlying the four broader groups above.

## 5.4 Organisation-level membership

```{r}
organisation_membership_total |>
  knitr::kable(
    col.names = c("Metric", "Organisations"),
    align = c("l", "r")
  )

organisation_status_summary |>
  knitr::kable(
    col.names = c("Organization status", "Organisations"),
    align = c("l", "r")
  )

organisation_type_summary |>
  knitr::kable(
    col.names = c("Organization type", "Organisations"),
    align = c("l", "r")
  )

organisation_membership |>
  slice_head(n = 10) |>
  knitr::kable(align = c("l", "l", "l", "l"))
```

The summary tables count all organisations, first as a total and then by
`Organization status` and `Organization type`. The two breakdowns describe
different attributes and should not be added together. The detail table
displays the first 10 organisations in alphabetical order as a preview; the
complete organisation-level dataset is saved as a CSV file. The complete table
has one row per non-blank standardised `school_name_key`.
`Organization name` is the most frequently occurring non-blank `company_name`
for that key. `Organization status` is Active if at least one associated
contact has `membership_status_clean` equal to Active; otherwise it is Not
active, including organisations with no recorded membership. `Organization
plan` lists the distinct school or organisation group plans among its contacts,
or No group membership plan when none is recorded. `Organization type` is
School when a school group plan is present, Organisation when an Organisation -
Group Membership plan is present without a school group plan, and Unknown when
neither type can be established.

<!--
# 6 Recommendations

## 6.1 High impact and low effort

```{r}
#| include: false
crm_recommendations <- tribble(
  ~priority, ~recommendation, ~evidence, ~proposed_action,
  "High impact / low effort",
  "Resolve technical data errors",
  "Four invalid non-missing postcodes and three parser-affected records.",
  "Review the seven restricted records and correct confirmed source errors.",
  "High impact / low effort",
  "Review membership conflicts",
  "Eleven status-tag disagreements, eleven multi-tag contacts and 49 Active records with past end dates.",
  "Confirm the intended status and remove or update obsolete membership tags.",
  "High impact / low effort",
  "Complete organisation names for group contacts",
  "183 Group Admin or Group Member contacts have no company name.",
  "Assign a verified school or organisation where one can be established.",
  "High impact / low effort",
  "Confirm required fields by contact type",
  "Thirteen selected fields are below 50% complete, but some blanks are legitimate.",
  "Define required fields separately for members, non-members and group contacts.",
  "Medium impact",
  "Create an approved school-name mapping",
  "156 basic variant groups and 489 fuzzy candidate pairs require review.",
  "Approve a source-name to standard-name table without overwriting raw values.",
  "Medium impact",
  "Review group-role coverage",
  "Among group-tagged organisations, 46 have no Admin, 15 have multiple Admins and 136 have no Member.",
  "Confirm whether each case is valid and update group-role tags where needed.",
  "Medium impact",
  "Reconcile missing status or tag categories",
  "Ninety contacts have a selected status without a tag and 32 have a tag without a comparable status.",
  "Document the authoritative membership field and synchronisation rule.",
  "Long-term opportunity",
  "Add controlled validation at data entry",
  "School names, postcodes, group roles and membership fields currently require retrospective checks.",
  "Use controlled lists and validation rules in the CRM input process.",
  "Long-term opportunity",
  "Automate recurring CRM quality monitoring",
  "The current reports provide repeatable checks for future exports.",
  "Schedule periodic refreshes and review changes in key quality measures.",
  "Long-term opportunity",
  "Collect dated engagement activity only if required",
  "The selected variables cannot measure active or dormant relationships.",
  "If STAV needs this measure, define qualifying interactions and capture a reliable last-activity date."
)

crm_recommendations |>
  filter(priority == "High impact / low effort") |>
  select(recommendation, evidence, proposed_action) |>
  knitr::kable(
    col.names = c("Recommendation", "Evidence", "Proposed action"),
    align = c("l", "l", "l")
  )
```

Each recommendation is based on the measured findings in Reports 02, 03 and 04.
These actions combine a material data-quality issue
with a relatively contained review or correction effort. Evidence gives the
current measured scope; the proposed action still requires confirmation before
source CRM records are changed.

## 6.2 Medium impact

```{r}
#| include: false
crm_recommendations |>
  filter(priority == "Medium impact") |>
  select(recommendation, evidence, proposed_action) |>
  knitr::kable(
    col.names = c("Recommendation", "Evidence", "Proposed action"),
    align = c("l", "l", "l")
  )
```

Each row links a measured issue to a proposed action. These actions require broader business review,
mapping or process decisions. They should follow the immediate technical and
record-level corrections.

## 6.3 Long-term opportunities

```{r}
#| include: false
crm_recommendations |>
  filter(priority == "Long-term opportunity") |>
  select(recommendation, evidence, proposed_action) |>
  knitr::kable(
    col.names = c("Recommendation", "Evidence", "Proposed action"),
    align = c("l", "l", "l")
  )
```

Each row is based on a recurring limitation identified in the current pipeline.
These are process and system improvements rather
than corrections to the current extract. They require ownership, implementation
planning and an agreed recurring workflow.
-->

```{r}
#| include: false
output_dir <- here("data", "analysis", "CRM", "06_summary")
dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)

write_csv(
  crm_health_check_status,
  file.path(output_dir, "crm_health_check_status.csv"),
  na = ""
)

write_csv(
  crm_recommendations,
  file.path(output_dir, "crm_recommendations.csv"),
  na = ""
)
```

# Conclusion

The current CRM export supports a repeatable audit of identifiers,
completeness, group-role coverage and membership consistency. The immediate
priority is to resolve the small number of technical errors, review membership
conflicts, complete missing organisations for group contacts and establish an
approved school-name mapping. Active and dormant relationship analysis has been
excluded because the available data cannot support a defensible classification.
