---
title: "STAV CRM Data Audit and Cleaning"
subtitle: "Contact, consent and membership lifecycle review"
format:
  html:
    toc: true
    number-sections: true
editor: visual
---

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(lubridate)
library(janitor)
library(here)
```

## Scope and privacy boundary

This report audits the CRM export `CRM20260729.csv`. The source contains direct identifiers and sensitive contact information. Individual names, email addresses, telephone numbers, street addresses, IP addresses and free-text notes are used only for validation and are not printed in this report or included in the analysis-ready output.

The raw export remains unchanged in `data/raw/CRM/`.

## Data import and CSV structure validation

All 71 source fields are initially imported as character values. This prevents R from silently selecting inconsistent types before the field rules are known.

```{r}
#| label: import-raw-crm

source_file <- here("data", "raw", "CRM", "CRM20260729.csv")
extract_date <- as.Date("2026-07-29")

crm_import <- suppressWarnings(
  read_csv(
    source_file,
    col_types = cols(.default = col_character()),
    trim_ws = FALSE,
    name_repair = "minimal",
    show_col_types = FALSE,
    progress = FALSE,
    lazy = FALSE
  )
)

# Capture parser problems before subsequent transformations remove the
# problems attribute.
crm_parse_problems <- problems(crm_import)
problem_data_rows <- sort(unique(crm_parse_problems$row - 1L))

original_names <- names(crm_import)
clean_names_vector <- make_clean_names(original_names)
names(crm_import) <- clean_names_vector

crm_import <- crm_import |>
  mutate(
    source_record_number = row_number(),
    structure_issue = source_record_number %in% problem_data_rows,
    across(
      all_of(setdiff(names(crm_import), "source_record_number")),
      ~ na_if(str_squish(.x), "")
    )
  )

crm_structure <- tibble(
  logical_contact_records = nrow(crm_import),
  source_fields = length(original_names),
  readr_parser_problem_rows = length(problem_data_rows),
  exact_duplicate_rows = sum(duplicated(crm_import |> select(-source_record_number)))
)

crm_structure
```

The parser problems originate in unquoted comma/JSON content in `notes`. Fields before `notes` remain available, but fields from `notes` onward are not trusted for the affected rows. The affected rows are retained and flagged.

```{r}
#| label: quarantine-structural-issues

post_notes_fields <- names(crm_import)[
  match("notes", names(crm_import)):match("company_department", names(crm_import))
]

crm_work <- crm_import |>
  mutate(
    across(
      all_of(post_notes_fields),
      ~ if_else(structure_issue, NA_character_, .x)
    )
  )
```

## Record grain and identifier validation

The audit treats each row as one CRM contact record. `ID` is assessed as the source-system primary key; email and `user_id` are tested as possible secondary keys but are not assumed to have the same meaning.

```{r}
#| label: identifier-validation

normalise_email <- function(x) str_to_lower(str_squish(x))
normalise_phone <- function(x) str_replace_all(x, "[^0-9+]", "")

identifier_validation <- tribble(
  ~identifier, ~non_missing, ~distinct_non_missing, ~duplicate_records, ~interpretation,
  "id", sum(!is.na(crm_work$id)), n_distinct(crm_work$id, na.rm = TRUE), sum(duplicated(crm_work$id) & !is.na(crm_work$id)), "Preferred CRM contact key",
  "email", sum(!is.na(crm_work$email)), n_distinct(normalise_email(crm_work$email), na.rm = TRUE), sum(duplicated(normalise_email(crm_work$email)) & !is.na(crm_work$email)), "Unique in this extract but personal and changeable",
  "user_id", sum(!is.na(crm_work$user_id)), n_distinct(crm_work$user_id, na.rm = TRUE), sum(duplicated(crm_work$user_id) & !is.na(crm_work$user_id)), "Not a contact primary key",
  "mobile_phone", sum(!is.na(crm_work$mobile_phone)), n_distinct(normalise_phone(crm_work$mobile_phone), na.rm = TRUE), sum(duplicated(normalise_phone(crm_work$mobile_phone)) & !is.na(crm_work$mobile_phone)), "Shared/reused; not a key",
  "primary_phone", sum(!is.na(crm_work$primary_phone)), n_distinct(normalise_phone(crm_work$primary_phone), na.rm = TRUE), sum(duplicated(normalise_phone(crm_work$primary_phone)) & !is.na(crm_work$primary_phone)), "Shared/reused; not a key"
)

identifier_validation
```

## Field groups and data dictionary

```{r}
#| label: data-dictionary

field_group <- function(field) case_when(
  field %in% c("id", "full_name", "first_name", "last_name", "email", "user_id", "owner_id") ~ "Identity",
  field %in% c("mobile_phone", "primary_phone", "primary_phone_extension", "street_address_1", "street_address_2", "city", "postal_zip", "region", "country", "time_zone", "ip_address") ~ "Contact and location",
  field %in% c("date_created", "birthday") ~ "Contact dates",
  field %in% c("optin_status", "date_optin_status_changed", "terms_agreement", "gdpr_consent", "gdpr_consent_date", "marketing_consent", "marketing_consent_date", "unsub_date", "unsub_reason", "unsub_feedback") ~ "Consent and subscription",
  field %in% c("lead_source", "source_page", "utm_campaign", "utm_content", "utm_medium", "utm_term", "utm_source") ~ "Acquisition",
  field %in% c("membership_status", "membership_plan", "subscription_start_date", "subscription_end_date", "date_of_cancellation", "reason_for_cancellation", "if_other", "date_retired") ~ "Membership lifecycle",
  field %in% c("year_level", "subjects", "job_title_position", "do_you_hold_an_identified_leadership_role", "do_you_hold_an_identified_leadership_role_other", "primary_teachers_and_schools", "secondary_teachers_and_schools", "p_12_teachers_tertiary_educators_academics_lab_technicians_and_p_12_schools_and_organisations", "your_role") ~ "Professional profile",
  field %in% c("publications_medium", "send_me_all_updates", "join_the_asta_mailing_list", "send_me_specific_updates") ~ "Communication preferences",
  field %in% c("last_sponsorship_purchase", "sponsorships_details_sections", "sponsorship_details_divisions", "supporting_sponsorships") ~ "Sponsorship",
  field %in% c("company_name", "company_website", "company_address", "company_phone", "company_phone_extension", "job_title", "company_department") ~ "Organisation",
  field %in% c("notes", "tags") ~ "Free text and tags",
  TRUE ~ "Other"
)

sensitivity <- function(field) case_when(
  field %in% c("full_name", "first_name", "last_name", "email", "birthday", "mobile_phone", "primary_phone", "street_address_1", "street_address_2", "postal_zip", "ip_address", "notes", "unsub_feedback") ~ "Direct/high-risk personal information",
  field %in% c("id", "user_id", "city", "region", "company_name", "company_address", "company_phone") ~ "Indirect identifier",
  TRUE ~ "Business/analytical field"
)

crm_data_dictionary <- tibble(
  source_name = original_names,
  field = clean_names_vector
) |>
  mutate(
    field_group = map_chr(field, field_group),
    sensitivity = map_chr(field, sensitivity),
    records = nrow(crm_work),
    raw_missing = map_int(field, ~ sum(is.na(crm_work[[.x]]))),
    raw_missing_rate = raw_missing / records,
    distinct_non_missing = map_int(field, ~ n_distinct(crm_work[[.x]], na.rm = TRUE))
  )

crm_data_dictionary
```

## Field-specific placeholder cleaning

`No` is not cleaned globally. It is retained in genuine yes/no consent fields and converted to missing only in fields where it is a system placeholder.

```{r}
#| label: placeholder-cleaning

no_placeholder_fields <- c(
  "birthday", "mobile_phone", "primary_phone", "primary_phone_extension",
  "street_address_1", "street_address_2", "city", "postal_zip", "region",
  "country", "time_zone", "ip_address", "lead_source", "source_page",
  "gdpr_consent_date", "marketing_consent_date", "utm_campaign",
  "utm_content", "utm_medium", "utm_term", "utm_source",
  "company_name", "company_website", "company_address", "company_phone",
  "company_phone_extension", "job_title", "company_department"
)

no_to_na <- function(x) if_else(str_to_lower(x) == "no", NA_character_, x)

crm_work <- crm_work |>
  mutate(across(all_of(no_placeholder_fields), no_to_na))

yes_no_to_logical <- function(x) case_when(
  str_to_lower(x) == "yes" ~ TRUE,
  str_to_lower(x) == "no" ~ FALSE,
  is.na(x) ~ NA,
  TRUE ~ NA
)

crm_work <- crm_work |>
  mutate(
    terms_agreement_clean = yes_no_to_logical(terms_agreement),
    gdpr_consent_clean = yes_no_to_logical(gdpr_consent),
    marketing_consent_clean = yes_no_to_logical(marketing_consent),
    email_clean = normalise_email(email),
    email_format_valid = str_detect(email_clean, "^[^[:space:]@]+@[^[:space:]@]+\\.[^[:space:]@]+$")
  )
```

## Date cleaning and validation

```{r}
#| label: date-cleaning

parse_crm_date <- function(x) {
  as.Date(suppressWarnings(parse_date_time(
    x,
    orders = c("ymd HMS", "ymd HM", "ymd", "dmy HMS", "dmy HM", "dmy", "mdy HMS", "mdy HM", "mdy"),
    tz = "Australia/Melbourne",
    quiet = TRUE
  )))
}

crm_work <- crm_work |>
  mutate(
    date_created_clean = parse_crm_date(date_created),
    date_optin_status_changed_clean = parse_crm_date(date_optin_status_changed),
    birthday_clean = parse_crm_date(birthday),
    gdpr_consent_date_clean = parse_crm_date(gdpr_consent_date),
    marketing_consent_date_clean = parse_crm_date(marketing_consent_date),
    unsub_date_clean = parse_crm_date(unsub_date),
    subscription_start_date_clean = parse_crm_date(subscription_start_date),
    subscription_end_date_clean = parse_crm_date(subscription_end_date),
    date_of_cancellation_clean = parse_crm_date(date_of_cancellation),
    date_retired_clean = parse_crm_date(date_retired),
    last_sponsorship_purchase_clean = parse_crm_date(last_sponsorship_purchase)
  )
```

```{r}
#| label: date-validation

date_fields_raw <- c(
  "date_created", "date_optin_status_changed", "birthday",
  "gdpr_consent_date", "marketing_consent_date", "unsub_date",
  "subscription_start_date", "subscription_end_date",
  "date_of_cancellation", "date_retired", "last_sponsorship_purchase"
)

date_fields_clean <- paste0(date_fields_raw, "_clean")

date_validation <- map2_dfr(date_fields_raw, date_fields_clean, function(raw, clean) {
  tibble(
    field = raw,
    non_missing_after_placeholder_cleaning = sum(!is.na(crm_work[[raw]])),
    successfully_parsed = sum(!is.na(crm_work[[clean]])),
    parse_failures = sum(!is.na(crm_work[[raw]]) & is.na(crm_work[[clean]])),
    minimum_date = if (all(is.na(crm_work[[clean]]))) as.Date(NA) else min(crm_work[[clean]], na.rm = TRUE),
    maximum_date = if (all(is.na(crm_work[[clean]]))) as.Date(NA) else max(crm_work[[clean]], na.rm = TRUE),
    future_after_extract = sum(crm_work[[clean]] > extract_date, na.rm = TRUE)
  )
})

date_validation
```

## Membership status and plan standardisation

```{r}
#| label: membership-cleaning

normalise_status <- function(x) {
  key <- x |>
    str_to_lower() |>
    str_squish() |>
    str_replace_all("[ _]", "-")
  case_when(
    key == "active" ~ "active",
    key == "on-hold" ~ "on-hold",
    key == "expired" ~ "expired",
    key == "cancelled" ~ "cancelled",
    key == "pending-cancel" ~ "pending-cancel",
    key == "paused" ~ "paused",
    key == "pending" ~ "pending",
    is.na(key) ~ NA_character_,
    TRUE ~ "review-unmapped"
  )
}

normalise_plan <- function(x) x |>
  str_replace_all("[–—]", "-") |>
  str_replace_all("\\s*-\\s*", " - ") |>
  str_squish()

crm_work <- crm_work |>
  mutate(
    membership_status_clean = normalise_status(membership_status),
    membership_plan_clean = normalise_plan(membership_plan),
    has_membership_record = !is.na(membership_status_clean) | !is.na(membership_plan_clean),
    membership_end_before_start = !is.na(subscription_start_date_clean) & !is.na(subscription_end_date_clean) & subscription_end_date_clean < subscription_start_date_clean,
    cancellation_before_start = !is.na(subscription_start_date_clean) & !is.na(date_of_cancellation_clean) & date_of_cancellation_clean < subscription_start_date_clean,
    active_with_past_end_date = membership_status_clean == "active" & !is.na(subscription_end_date_clean) & subscription_end_date_clean < extract_date,
    membership_status_without_plan = !is.na(membership_status_clean) & is.na(membership_plan_clean),
    consent_date_without_consent_value = (!is.na(marketing_consent_date_clean) & is.na(marketing_consent_clean)) | (!is.na(gdpr_consent_date_clean) & is.na(gdpr_consent_clean)),
    unsubscribed_without_reason = !is.na(unsub_date_clean) & is.na(unsub_reason)
  )
```

```{r}
#| label: membership-summary

membership_status_summary <- crm_work |>
  filter(!structure_issue) |>
  count(membership_status_clean, sort = TRUE, name = "contact_records")

membership_plan_summary <- crm_work |>
  filter(!structure_issue, !is.na(membership_plan_clean)) |>
  count(membership_plan_clean, sort = TRUE, name = "contact_records")

membership_status_summary
membership_plan_summary
```

## Consolidated validation findings

```{r}
#| label: consolidated-validation

crm_validation <- tibble(
  check = c(
    "Contact records",
    "Source fields",
    "Parser/structure issue records",
    "Missing CRM ID",
    "Duplicate CRM ID",
    "Missing email",
    "Duplicate normalised email",
    "Invalid email format",
    "Membership end before start",
    "Cancellation before membership start",
    "Active status with past end date",
    "Membership status without plan",
    "Consent date without consent value",
    "Unsubscribe date without reason"
  ),
  result = c(
    nrow(crm_work),
    length(original_names),
    sum(crm_work$structure_issue),
    sum(is.na(crm_work$id)),
    sum(duplicated(crm_work$id) & !is.na(crm_work$id)),
    sum(is.na(crm_work$email_clean)),
    sum(duplicated(crm_work$email_clean) & !is.na(crm_work$email_clean)),
    sum(!crm_work$email_format_valid, na.rm = TRUE),
    sum(crm_work$membership_end_before_start, na.rm = TRUE),
    sum(crm_work$cancellation_before_start, na.rm = TRUE),
    sum(crm_work$active_with_past_end_date, na.rm = TRUE),
    sum(crm_work$membership_status_without_plan, na.rm = TRUE),
    sum(crm_work$consent_date_without_consent_value, na.rm = TRUE),
    sum(crm_work$unsubscribed_without_reason, na.rm = TRUE)
  )
)

crm_validation
```

## Analysis-ready dataset

The output retains the CRM ID for internal joins but excludes direct personal information and free-text notes. `structure_issue` remains visible so affected records are never silently treated as fully valid.

```{r}
#| label: analysis-ready-data

crm_contacts_clean <- crm_work |>
  transmute(
    crm_record_id = id,
    user_id,
    source_record_number,
    structure_issue,
    date_created = date_created_clean,
    optin_status_code = suppressWarnings(as.integer(optin_status)),
    terms_agreement = terms_agreement_clean,
    gdpr_consent = gdpr_consent_clean,
    gdpr_consent_date = gdpr_consent_date_clean,
    marketing_consent = marketing_consent_clean,
    marketing_consent_date = marketing_consent_date_clean,
    unsub_date = unsub_date_clean,
    unsub_reason,
    lead_source,
    source_page,
    utm_source,
    utm_medium,
    utm_campaign,
    utm_content,
    utm_term,
    membership_status = membership_status_clean,
    membership_plan = membership_plan_clean,
    subscription_start_date = subscription_start_date_clean,
    subscription_end_date = subscription_end_date_clean,
    cancellation_date = date_of_cancellation_clean,
    cancellation_reason = reason_for_cancellation,
    retirement_date = date_retired_clean,
    year_level,
    subjects,
    professional_role = your_role,
    job_title_position,
    leadership_role = do_you_hold_an_identified_leadership_role,
    company_name,
    company_website,
    company_job_title = job_title,
    company_department,
    has_membership_record,
    membership_end_before_start,
    cancellation_before_start,
    active_with_past_end_date,
    membership_status_without_plan,
    consent_date_without_consent_value,
    unsubscribed_without_reason
  )

crm_contacts_clean
```

## Cleaning log and outputs

```{r}
#| label: cleaning-log

crm_cleaning_log <- tribble(
  ~step, ~area, ~rule, ~status, ~records_affected, ~reason,
  1L, "Raw import", "Import all 71 fields as character and preserve raw file", "Completed", nrow(crm_import), "Avoid premature type conversion",
  2L, "CSV structure", "Capture readr parser problems and flag affected records", "Completed", sum(crm_import$structure_issue), "Prevent post-notes field misalignment from entering analysis",
  3L, "Column names", "Convert source headers to snake_case", "Completed", length(original_names), "Consistent R field names",
  4L, "Whitespace", "Trim leading, trailing and repeated whitespace; blank strings become NA", "Completed", NA_integer_, "Consistent missing-value handling",
  5L, "No placeholders", "Convert No to NA only in explicitly listed non-boolean fields", "Completed", NA_integer_, "Preserve genuine consent No values",
  6L, "Identifiers", "Use CRM ID as contact key; retain user_id only as a secondary field", "Completed", nrow(crm_work), "ID is complete and unique in this extract",
  7L, "Consent", "Normalise yes/no values to logical fields", "Completed", NA_integer_, "Auditable consent validation",
  8L, "Dates", "Parse supported CRM date formats; retain parse-failure counts", "Completed", NA_integer_, "Consistent lifecycle dates",
  9L, "Membership status", "Normalise case, spaces and hyphens to controlled labels", "Completed", sum(!is.na(crm_work$membership_status_clean)), "Consistent membership reporting",
  10L, "Membership plan", "Normalise dash characters and spacing without merging plan types", "Completed", sum(!is.na(crm_work$membership_plan_clean)), "Remove typography-only variants",
  11L, "Privacy", "Exclude names, email, phones, addresses, IP and notes from analysis-ready output", "Completed", nrow(crm_work), "Data minimisation",
  12L, "Quality flags", "Retain status/date/consent inconsistencies as flags; do not delete records", "Completed", NA_integer_, "Support review without unsupported correction"
)

crm_cleaning_log
```

```{r}
#| label: export-clean-outputs

crm_output_dir <- here("data", "clean", "CRM")
dir.create(crm_output_dir, recursive = TRUE, showWarnings = FALSE)

write_csv(crm_contacts_clean, file.path(crm_output_dir, "crm_contacts_clean.csv"), na = "")
write_csv(crm_data_dictionary, file.path(crm_output_dir, "crm_data_dictionary.csv"), na = "")
write_csv(crm_validation, file.path(crm_output_dir, "crm_validation_summary.csv"), na = "")
write_csv(crm_cleaning_log, file.path(crm_output_dir, "crm_cleaning_log.csv"), na = "")

crm_output_manifest <- tibble(
  output_file = c(
    "crm_contacts_clean.csv", "crm_data_dictionary.csv",
    "crm_validation_summary.csv", "crm_cleaning_log.csv"
  ),
  purpose = c(
    "De-identified analysis-ready contact and membership records",
    "Field inventory, grouping, sensitivity and raw completeness",
    "Key structural, identifier, consent and membership checks",
    "Documented cleaning and validation rules"
  ),
  rows = c(
    nrow(crm_contacts_clean), nrow(crm_data_dictionary),
    nrow(crm_validation), nrow(crm_cleaning_log)
  ),
  generated_from = basename(source_file)
)

write_csv(crm_output_manifest, file.path(crm_output_dir, "crm_output_manifest.csv"), na = "")
crm_output_manifest
```

## Initial audit interpretation

- `ID` is the appropriate CRM contact key for this extract.
- Email is complete and unique but is personal information and is not exported to the analysis-ready file.
- `user_id` is not a contact primary key and should not be used to deduplicate contacts.
- Placeholder `No` values require field-specific treatment.
- Parser-affected records are retained and transparently flagged.
- Membership and consent inconsistencies are flags for review, not reasons to delete contacts or overwrite source values.

The next analytical layer should examine member composition, membership plan mix, lifecycle status, consent coverage and acquisition-field completeness using the validated clean output.
