---
title: "STAV GA4 Website Analytics"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
library(janitor)
```

## 1. Data Import

The GA4 exports contain nine metadata rows before the column headings. These
rows are skipped when the files are imported.

```{r}
demographic <- read_csv(
  here("data", "raw", "GA4", "Demographic_details_Country.csv"),
  skip = 9,
  show_col_types = FALSE
)

events <- read_csv(
  here("data", "raw", "GA4", "Events_Event_name.csv"),
  skip = 9,
  show_col_types = FALSE
)

landing_page <- read_csv(
  here("data", "raw", "GA4", "Landing_page_Landing_page.csv"),
  skip = 9,
  show_col_types = FALSE
)

pages <- read_csv(
  here("data", "raw", "GA4", "Pages_and_screens_Page_path_and_screen_class.csv"),
  skip = 9,
  show_col_types = FALSE
)

traffic <- read_csv(
  here(
    "data", "raw", "GA4",
    "Traffic_acquisition_Session_primary_channel_group_(Default_Channel_Group).csv"
  ),
  skip = 9,
  show_col_types = FALSE
)
```

## 2. Data Structure

Standardise all column names to snake case before carrying out validation or
analysis.

```{r}
demographic <- demographic |> clean_names()
events <- events |> clean_names()
landing_page <- landing_page |> clean_names()
pages <- pages |> clean_names()
traffic <- traffic |> clean_names()
```

```{r}
ga4_structure <- tibble(
  dataset = c("demographic", "events", "landing_page", "pages", "traffic"),
  rows = c(
    nrow(demographic), nrow(events), nrow(landing_page),
    nrow(pages), nrow(traffic)
  ),
  columns = c(
    ncol(demographic), ncol(events), ncol(landing_page),
    ncol(pages), ncol(traffic)
  )
)

ga4_structure
```

```{r}
list(
  demographic = names(demographic),
  events = names(events),
  landing_page = names(landing_page),
  pages = names(pages),
  traffic = names(traffic)
)
```

## 3. Initial Data Validation

These checks report possible issues but do not remove or modify observations.

```{r}
ga4_data <- list(
  demographic = demographic,
  events = events,
  landing_page = landing_page,
  pages = pages,
  traffic = traffic
)

ga4_validation <- imap_dfr(
  ga4_data,
  ~ tibble(
    dataset = .y,
    rows = nrow(.x),
    columns = ncol(.x),
    missing_values = sum(is.na(.x)),
    duplicate_rows = sum(duplicated(.x))
  )
)

ga4_validation
```

## 4. Events Data Audit

Review the events currently collected in GA4 and assess whether meaningful
website conversions are being tracked.

Classify the recorded events into standard GA4 activity, content interactions
and potential business conversions. This classification supports the audit; it
does not alter the original event names or counts.

```{r}
events_audit <- events |>
  mutate(
    event_category = case_when(
      event_name %in% c(
        "page_view",
        "session_start",
        "first_visit",
        "user_engagement",
        "scroll"
      ) ~ "Automatically collected or enhanced measurement",

      str_detect(
        event_name,
        regex(
          "form|register|registration|sign_up|purchase|member",
          ignore_case = TRUE
        )
      ) ~ "Potential business conversion",

      str_detect(
        event_name,
        regex(
          "click|download|file|video",
          ignore_case = TRUE
        )
      ) ~ "Content interaction",

      TRUE ~ "Other"
    ),

    tracking_status = case_when(
      event_category == "Potential business conversion" ~
        "Review whether this should be a key event",

      event_category ==
        "Automatically collected or enhanced measurement" ~
        "Standard GA4 event",

      TRUE ~ "Review event definition"
    )
  ) |>
  arrange(desc(event_count))

events_audit |>
  select(
    event_name,
    event_count,
    total_users,
    event_category,
    tracking_status
  )
```

Search for events whose names suggest a registration, form submission,
membership action or other business conversion.

```{r}
potential_conversion_events <- events_audit |>
  filter(
    str_detect(
      event_name,
      regex(
        "form|register|registration|sign_up|purchase|member",
        ignore_case = TRUE
      )
    )
  )

potential_conversion_events
```

Summarise the recorded event activity by audit category. The user counts below
must not be interpreted as distinct users across categories because the same
person can trigger more than one event type.

```{r}
event_category_summary <- events_audit |>
  group_by(event_category) |>
  summarise(
    number_of_event_types = n(),
    total_event_count = sum(event_count, na.rm = TRUE),
    total_users_across_event_rows = sum(total_users, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(total_event_count))

event_category_summary
```

Record the principal data-quality checks for the events export.

```{r}
events_findings <- tibble(
  check = c(
    "Number of event types",
    "Missing event names",
    "Duplicate event names",
    "Potential conversion events",
    "Event types with zero counts"
  ),
  result = c(
    nrow(events),
    sum(is.na(events$event_name) | events$event_name == ""),
    sum(duplicated(events$event_name)),
    nrow(potential_conversion_events),
    sum(events$event_count == 0, na.rm = TRUE)
  )
)

events_findings
```

## 5. Traffic Acquisition Data Audit

Review the channels through which users reached the STAV website and assess
whether campaign traffic is being classified reliably.

```{r}
traffic_audit <- traffic |>
  rename(
    channel =
      session_primary_channel_group_default_channel_group
  ) |>
  mutate(
    session_share =
      sessions / sum(sessions, na.rm = TRUE),

    engagement_rate_check =
      engaged_sessions / sessions,

    engagement_rate_difference =
      abs(engagement_rate - engagement_rate_check),

    channel_flag = case_when(
      channel == "Direct" ~
        "Review possible untagged campaign traffic",

      channel == "Unassigned" ~
        "Traffic source could not be classified",

      channel == "Email" & sessions < 10 ~
        "Email traffic may be under-recorded",

      key_events == 0 ~
        "No key events recorded",

      TRUE ~
        "No immediate issue identified"
    )
  ) |>
  arrange(desc(sessions))

traffic_audit |>
  select(
    channel,
    sessions,
    session_share,
    engagement_rate,
    average_engagement_time_per_session,
    key_events,
    channel_flag
  )
```

Check the internal consistency and validity of the traffic-acquisition data.

```{r}
traffic_validation <- tibble(
  check = c(
    "Missing channel names",
    "Duplicate channel names",
    "Negative session values",
    "Engaged sessions greater than sessions",
    "Engagement rates outside 0-1",
    "Engagement-rate calculation differences",
    "Channels with zero key events"
  ),

  result = c(
    sum(
      is.na(traffic_audit$channel) |
        traffic_audit$channel == ""
    ),

    sum(duplicated(traffic_audit$channel)),

    sum(traffic_audit$sessions < 0, na.rm = TRUE),

    sum(
      traffic_audit$engaged_sessions >
        traffic_audit$sessions,
      na.rm = TRUE
    ),

    sum(
      traffic_audit$engagement_rate < 0 |
        traffic_audit$engagement_rate > 1,
      na.rm = TRUE
    ),

    sum(
      traffic_audit$engagement_rate_difference >
        0.000001,
      na.rm = TRUE
    ),

    sum(traffic_audit$key_events == 0, na.rm = TRUE)
  )
)

traffic_validation
```

Calculate overall traffic totals. The overall engagement rate is weighted by
sessions rather than calculated as a simple average of channel rates.

```{r}
traffic_summary <- traffic_audit |>
  summarise(
    total_sessions =
      sum(sessions, na.rm = TRUE),

    total_engaged_sessions =
      sum(engaged_sessions, na.rm = TRUE),

    overall_engagement_rate =
      total_engaged_sessions / total_sessions,

    total_event_count =
      sum(event_count, na.rm = TRUE),

    total_key_events =
      sum(key_events, na.rm = TRUE)
  )

traffic_summary
```

Quantify traffic categories that may indicate incomplete campaign tagging.
These results identify areas for investigation; they do not prove that Direct
traffic originated from a particular campaign.

```{r}
tracking_risk_summary <- traffic_audit |>
  summarise(
    direct_sessions =
      sum(sessions[channel == "Direct"], na.rm = TRUE),

    direct_share =
      sum(session_share[channel == "Direct"], na.rm = TRUE),

    unassigned_sessions =
      sum(sessions[channel == "Unassigned"], na.rm = TRUE),

    unassigned_share =
      sum(session_share[channel == "Unassigned"], na.rm = TRUE),

    email_sessions =
      sum(sessions[channel == "Email"], na.rm = TRUE),

    email_share =
      sum(session_share[channel == "Email"], na.rm = TRUE)
  )

tracking_risk_summary
```

## 6. Landing Page Data Audit

Review the pages through which sessions entered the STAV website and identify
path-formatting issues, unclassified landing pages and pages that may require
further investigation. The original landing-page field is retained unchanged.

```{r}
landing_page_audit <- landing_page |>
  mutate(
    landing_page_original = landing_page,

    has_query_parameter =
      str_detect(landing_page, fixed("?")),

    landing_page_standardised = landing_page |>
      str_squish() |>
      str_remove("\\?.*$"),

    landing_page_standardised = case_when(
      landing_page_standardised == "/" ~ "/",
      landing_page_standardised == "(not set)" ~ "(not set)",
      TRUE ~ str_remove(landing_page_standardised, "/+$")
    ),

    landing_page_flag = case_when(
      is.na(landing_page) | landing_page == "" ~
        "Missing landing-page value",

      landing_page == "(not set)" ~
        "Landing page was not identified",

      has_query_parameter ~
        "Query parameters removed in standardised path",

      landing_page != "/" & str_ends(landing_page, "/") ~
        "Trailing slash removed in standardised path",

      key_events == 0 ~
        "No key events recorded",

      TRUE ~
        "No immediate issue identified"
    )
  ) |>
  arrange(desc(sessions))

landing_page_audit |>
  select(
    landing_page_original,
    landing_page_standardised,
    sessions,
    active_users,
    average_engagement_time_per_session,
    key_events,
    landing_page_flag
  ) |>
  slice_head(n = 20)
```

Check the structural quality of the landing-page export. Duplicate
standardised paths indicate that multiple exported rows may refer to the same
page after query parameters and trailing slashes are removed.

```{r}
standardised_landing_page_duplicates <- landing_page_audit |>
  filter(landing_page_standardised != "(not set)") |>
  count(
    landing_page_standardised,
    name = "number_of_rows"
  ) |>
  filter(number_of_rows > 1) |>
  arrange(desc(number_of_rows))

landing_page_validation <- tibble(
  check = c(
    "Missing landing-page values",
    "Exact duplicate rows",
    "Duplicate original landing-page paths",
    "Rows recorded as (not set)",
    "Paths containing query parameters",
    "Non-root paths ending in a slash",
    "Duplicated standardised paths",
    "Rows with zero key events"
  ),

  result = c(
    sum(
      is.na(landing_page_audit$landing_page_original) |
        landing_page_audit$landing_page_original == ""
    ),

    sum(duplicated(landing_page)),

    sum(duplicated(landing_page$landing_page)),

    sum(
      landing_page_audit$landing_page_original == "(not set)",
      na.rm = TRUE
    ),

    sum(landing_page_audit$has_query_parameter, na.rm = TRUE),

    sum(
      landing_page_audit$landing_page_original != "/" &
        str_ends(landing_page_audit$landing_page_original, "/"),
      na.rm = TRUE
    ),

    nrow(standardised_landing_page_duplicates),

    sum(landing_page_audit$key_events == 0, na.rm = TRUE)
  )
)

landing_page_validation
```

Quantify the scale of unidentified landing-page traffic and confirm the
overall volume represented by this export.

```{r}
landing_page_summary <- landing_page_audit |>
  summarise(
    total_sessions = sum(sessions, na.rm = TRUE),

    total_active_users_across_rows =
      sum(active_users, na.rm = TRUE),

    not_set_sessions =
      sum(
        sessions[landing_page_original == "(not set)"],
        na.rm = TRUE
      ),

    not_set_session_share =
      not_set_sessions / total_sessions,

    total_key_events =
      sum(key_events, na.rm = TRUE)
  )

landing_page_summary
```

Identify frequently used landing pages with very short average engagement
time. These thresholds are screening rules for investigation rather than a
formal definition of poor performance.

```{r}
high_traffic_low_engagement_pages <- landing_page_audit |>
  filter(
    landing_page_original != "(not set)",
    sessions >= 100,
    average_engagement_time_per_session < 10
  ) |>
  select(
    landing_page_original,
    sessions,
    active_users,
    average_engagement_time_per_session,
    key_events
  ) |>
  arrange(
    average_engagement_time_per_session,
    desc(sessions)
  )

high_traffic_low_engagement_pages
```

Review standardised paths that correspond to more than one original exported
row before deciding whether they should be aggregated for later analysis.

```{r}
standardised_landing_page_duplicates |>
  slice_head(n = 20)
```

## 7. Pages and Screens Data Audit

Review page-level website activity and identify path-formatting issues,
possible tracking anomalies and content that may require further review. The
original page path is retained, and the engagement metric is interpreted as
average engagement time per active user rather than time per view.

```{r}
pages_audit <- pages |>
  mutate(
    page_path_original = page_path_and_screen_class,

    has_query_parameter =
      str_detect(page_path_original, fixed("?")),

    page_path_standardised = page_path_original |>
      str_squish() |>
      str_remove("\\?.*$"),

    page_path_standardised = case_when(
      page_path_standardised == "/" ~ "/",
      page_path_standardised == "(not set)" ~ "(not set)",
      TRUE ~ str_remove(page_path_standardised, "/+$")
    ),

    page_path_template = page_path_standardised |>
      str_replace_all(
        "(/view-order/)\\d+",
        "\\1{id}"
      ) |>
      str_replace_all(
        "(/view-subscription/)\\d+",
        "\\1{id}"
      ) |>
      str_replace_all(
        "(/members-area/)\\d+",
        "\\1{id}"
      ) |>
      str_replace_all(
        "(/groups/)\\d+",
        "\\1{id}"
      ) |>
      str_replace_all(
        "(/join-group/)[A-Za-z0-9_-]{16,}",
        "\\1{token}"
      ),

    page_path_was_templated =
      page_path_template != page_path_standardised,

    page_path_flag = case_when(
      is.na(page_path_original) | page_path_original == "" ~
        "Missing page path",

      page_path_original == "(not set)" ~
        "Page path was not identified",

      has_query_parameter ~
        "Query parameters removed in standardised path",

      page_path_original != "/" &
        str_ends(page_path_original, "/") ~
        "Trailing slash removed in standardised path",

      key_events == 0 ~
        "No key events recorded",

      TRUE ~
        "No immediate issue identified"
    )
  ) |>
  arrange(desc(views))

pages_audit |>
  select(
    page_path_original,
    page_path_standardised,
    page_path_template,
    views,
    active_users,
    views_per_active_user,
    average_engagement_time_per_active_user,
    key_events,
    page_path_flag
  ) |>
  slice_head(n = 20)
```

Check page-path completeness, duplication and formatting. A duplicated
standardised path means that multiple exported rows refer to the same path
after query parameters and trailing slashes are removed.

```{r}
standardised_page_path_duplicates <- pages_audit |>
  filter(page_path_standardised != "(not set)") |>
  count(
    page_path_standardised,
    name = "number_of_rows"
  ) |>
  filter(number_of_rows > 1) |>
  arrange(desc(number_of_rows))

pages_validation <- tibble(
  check = c(
    "Missing page-path values",
    "Exact duplicate rows",
    "Duplicate original page paths",
    "Rows recorded as (not set)",
    "Paths containing query parameters",
    "Non-root paths ending in a slash",
    "Duplicated standardised paths",
    "Rows with zero key events",
    "Rows with negative numeric values",
    "Rows with views below active users"
  ),

  result = c(
    sum(
      is.na(pages_audit$page_path_original) |
        pages_audit$page_path_original == ""
    ),

    sum(duplicated(pages)),

    sum(duplicated(pages$page_path_and_screen_class)),

    sum(
      pages_audit$page_path_original == "(not set)",
      na.rm = TRUE
    ),

    sum(pages_audit$has_query_parameter, na.rm = TRUE),

    sum(
      pages_audit$page_path_original != "/" &
        str_ends(pages_audit$page_path_original, "/"),
      na.rm = TRUE
    ),

    nrow(standardised_page_path_duplicates),

    sum(pages_audit$key_events == 0, na.rm = TRUE),

    sum(
      pages_audit |>
        select(where(is.numeric)) |>
        as.matrix() < 0,
      na.rm = TRUE
    ),

    sum(pages_audit$views < pages_audit$active_users, na.rm = TRUE)
  )
)

pages_validation
```

Reconcile total page views in the Pages and Screens export with the
`page_view` count in the Events export. Any difference is retained and
reported rather than manually adjusted.

```{r}
page_view_event_count <- events |>
  filter(event_name == "page_view") |>
  summarise(value = sum(event_count, na.rm = TRUE)) |>
  pull(value)

pages_summary <- pages_audit |>
  summarise(
    total_views = sum(views, na.rm = TRUE),

    page_view_event_count = page_view_event_count,

    view_count_difference =
      total_views - page_view_event_count,

    total_event_count_across_page_rows =
      sum(event_count, na.rm = TRUE),

    total_key_events =
      sum(key_events, na.rm = TRUE)
  )

pages_summary
```

Identify frequently viewed pages with very short average engagement time per
active user. These thresholds are screening rules and do not by themselves
establish that a page performs poorly.

```{r}
high_view_low_engagement_pages <- pages_audit |>
  filter(
    page_path_original != "(not set)",
    views >= 100,
    average_engagement_time_per_active_user < 10
  ) |>
  select(
    page_path_original,
    views,
    active_users,
    views_per_active_user,
    average_engagement_time_per_active_user,
    key_events
  ) |>
  arrange(
    average_engagement_time_per_active_user,
    desc(views)
  )

high_view_low_engagement_pages
```

Review pages whose paths suggest membership, account, login or subscription
activity. This is a path-based content grouping and does not identify users as
members.

```{r}
member_related_pages <- pages_audit |>
  filter(
    str_detect(
      page_path_original,
      regex(
        "my-account|member|membership|subscription|login|sts-resources",
        ignore_case = TRUE
      )
    )
  ) |>
  select(
    page_path_original,
    views,
    active_users,
    views_per_active_user,
    average_engagement_time_per_active_user,
    key_events
  ) |>
  arrange(desc(views))

member_related_pages |>
  slice_head(n = 30)
```

### Dynamic Member-Path Templates

For analytical reporting, replace confirmed dynamic member-area identifiers
with placeholders. This creates an additional analysis field and does not
change the original GA4 path or the website.

```{r}
page_template_validation <- pages_audit |>
  summarise(
    original_page_paths =
      n_distinct(page_path_original),

    standardised_page_paths =
      n_distinct(page_path_standardised),

    templated_page_paths =
      n_distinct(page_path_template),

    rows_changed_by_templating =
      sum(page_path_was_templated, na.rm = TRUE),

    distinct_standardised_paths_changed =
      n_distinct(
        page_path_standardised[page_path_was_templated]
      ),

    distinct_templates_created =
      n_distinct(
        page_path_template[page_path_was_templated]
      )
  )

page_template_validation
```

Review how the original member-area paths map to the new templates.

```{r}
member_path_template_mapping <- pages_audit |>
  filter(page_path_was_templated) |>
  select(
    page_path_original,
    page_path_standardised,
    page_path_template,
    views,
    event_count
  ) |>
  arrange(
    page_path_template,
    desc(views)
  )

member_path_template_mapping |>
  slice_head(n = 40)
```

Aggregate additive activity measures by template. Active users are not summed
because the same user may appear in more than one original path row.

```{r}
member_page_template_summary <- pages_audit |>
  filter(page_path_was_templated) |>
  group_by(page_path_template) |>
  summarise(
    original_path_rows = n(),
    total_views = sum(views, na.rm = TRUE),
    total_event_count = sum(event_count, na.rm = TRUE),
    total_key_events = sum(key_events, na.rm = TRUE),
    .groups = "drop"
  ) |>
  arrange(desc(total_views))

member_page_template_summary |>
  slice_head(n = 30)
```

Search for remaining member-area paths containing long tokens or numeric path
segments that were not changed by the confirmed templating rules. These are
review candidates, not values that should be automatically replaced.

```{r}
unresolved_dynamic_member_paths <- pages_audit |>
  filter(
    str_detect(
      page_path_standardised,
      regex("my-account|member|membership", ignore_case = TRUE)
    ),
    !page_path_was_templated,
    str_detect(
      page_path_standardised,
      "(/\\d{4,})(/|$)|(/[A-Fa-f0-9]{16,})(/|$)"
    )
  ) |>
  select(
    page_path_original,
    page_path_standardised,
    views,
    event_count
  ) |>
  arrange(desc(views))

unresolved_dynamic_member_paths |>
  slice_head(n = 30)
```

Review the most frequently duplicated standardised page paths before deciding
whether to aggregate them for later analysis.

```{r}
standardised_page_path_duplicates |>
  slice_head(n = 30)
```

## 8. Demographic Data Audit

Review the geographic distribution of GA4 activity and identify countries
whose volume and engagement patterns require further investigation. Review
flags are screening rules only and do not establish that traffic is automated,
invalid or irrelevant.

```{r}
demographic_audit <- demographic |>
  mutate(
    calculated_engaged_sessions_per_active_user =
      engaged_sessions / active_users,

    engaged_sessions_per_user_difference = abs(
      engaged_sessions_per_active_user -
        calculated_engaged_sessions_per_active_user
    ),

    new_user_ratio =
      new_users / active_users,

    low_engagement_rate_flag =
      active_users >= 500 & engagement_rate < 0.15,

    short_engagement_time_flag =
      active_users >= 500 &
        average_engagement_time_per_active_user < 10,

    review_flag =
      low_engagement_rate_flag |
        short_engagement_time_flag,

    review_reason = case_when(
      low_engagement_rate_flag & short_engagement_time_flag ~
        "High-volume country with low engagement rate and short engagement time",

      low_engagement_rate_flag ~
        "High-volume country with low engagement rate",

      short_engagement_time_flag ~
        "High-volume country with short engagement time",

      TRUE ~
        "No review flag under current screening rules"
    )
  ) |>
  arrange(desc(active_users))

demographic_audit |>
  select(
    country,
    active_users,
    new_users,
    engaged_sessions,
    engagement_rate,
    average_engagement_time_per_active_user,
    new_user_ratio,
    review_flag,
    review_reason
  ) |>
  slice_head(n = 20)
```

Check completeness, uniqueness, numeric validity and the internal calculation
of engaged sessions per active user. New users can occasionally exceed active
users because the GA4 metrics are defined differently, so these rows are
reported for interpretation rather than treated as invalid records.

```{r}
demographic_validation <- tibble(
  check = c(
    "Missing country values",
    "Duplicate country names",
    "Rows with negative numeric values",
    "Engagement rates outside 0-1",
    "Engaged-sessions-per-user calculation differences",
    "Rows with new users above active users",
    "Countries meeting review rules"
  ),

  result = c(
    sum(
      is.na(demographic_audit$country) |
        demographic_audit$country == ""
    ),

    sum(duplicated(demographic_audit$country)),

    sum(
      demographic_audit |>
        select(
          active_users,
          new_users,
          engaged_sessions,
          engagement_rate,
          engaged_sessions_per_active_user,
          average_engagement_time_per_active_user,
          event_count,
          key_events,
          user_key_event_rate,
          total_revenue
        ) |>
        as.matrix() < 0,
      na.rm = TRUE
    ),

    sum(
      demographic_audit$engagement_rate < 0 |
        demographic_audit$engagement_rate > 1,
      na.rm = TRUE
    ),

    sum(
      demographic_audit$
        engaged_sessions_per_user_difference > 0.000001,
      na.rm = TRUE
    ),

    sum(
      demographic_audit$new_users >
        demographic_audit$active_users,
      na.rm = TRUE
    ),

    sum(demographic_audit$review_flag, na.rm = TRUE)
  )
)

demographic_validation
```

Review all countries selected by the transparent screening rules. These rows
remain in the data and require source, medium, hostname and landing-page
evidence before any exclusion decision could be justified.

```{r}
demographic_review_countries <- demographic_audit |>
  filter(review_flag) |>
  select(
    country,
    active_users,
    new_users,
    engaged_sessions,
    engagement_rate,
    average_engagement_time_per_active_user,
    event_count,
    new_user_ratio,
    review_reason
  ) |>
  arrange(desc(active_users))

demographic_review_countries
```

Compare reported activity associated with Australia and overseas countries.
Country-level active-user and new-user values are additive metrics in this
export but may double-count people who were recorded in more than one country;
the summed fields are labelled accordingly.

```{r}
demographic_market <- demographic_audit |>
  mutate(
    market = if_else(
      country == "Australia",
      "Australia",
      "Overseas"
    )
  ) |>
  group_by(market) |>
  summarise(
    number_of_country_rows = n(),

    reported_active_users_sum =
      sum(active_users, na.rm = TRUE),

    reported_new_users_sum =
      sum(new_users, na.rm = TRUE),

    engaged_sessions =
      sum(engaged_sessions, na.rm = TRUE),

    event_count =
      sum(event_count, na.rm = TRUE),

    engaged_sessions_per_reported_active_user =
      engaged_sessions / reported_active_users_sum,

    events_per_reported_active_user =
      event_count / reported_active_users_sum,

    review_country_rows =
      sum(review_flag, na.rm = TRUE),

    .groups = "drop"
  )

demographic_market
```

Assess how much the country rows selected for review affect additive activity
metrics. The comparison does not delete data and does not produce an official
GA4 user total or engagement rate.

```{r}
demographic_sensitivity <- bind_rows(
  demographic_audit |>
    summarise(
      scenario = "All country rows",
      country_rows = n(),
      reported_active_users_sum =
        sum(active_users, na.rm = TRUE),
      engaged_sessions =
        sum(engaged_sessions, na.rm = TRUE),
      event_count =
        sum(event_count, na.rm = TRUE)
    ),

  demographic_audit |>
    filter(!review_flag) |>
    summarise(
      scenario = "Excluding review-flagged country rows",
      country_rows = n(),
      reported_active_users_sum =
        sum(active_users, na.rm = TRUE),
      engaged_sessions =
        sum(engaged_sessions, na.rm = TRUE),
      event_count =
        sum(event_count, na.rm = TRUE)
    )
) |>
  mutate(
    engaged_sessions_per_reported_active_user =
      engaged_sessions / reported_active_users_sum,

    events_per_reported_active_user =
      event_count / reported_active_users_sum
  )

demographic_sensitivity
```

Confirm whether all country rows also report zero Key Events and revenue.

```{r}
demographic_key_event_summary <- demographic_audit |>
  summarise(
    country_rows = n(),
    rows_with_zero_key_events =
      sum(key_events == 0, na.rm = TRUE),
    total_key_events =
      sum(key_events, na.rm = TRUE),
    total_revenue =
      sum(total_revenue, na.rm = TRUE)
  )

demographic_key_event_summary
```

## 9. Clean Data Outputs and Cleaning Log

Create analysis-ready datasets without modifying the source exports. Each
output retains the original business dimension, adds documented audit fields
where relevant, and records the common analysis period.

```{r}
analysis_start_date <- as.Date("2025-07-01")
analysis_end_date <- as.Date("2026-06-30")

demographic_clean <- demographic_audit |>
  transmute(
    country,
    active_users,
    new_users,
    engaged_sessions,
    engagement_rate,
    engaged_sessions_per_active_user,
    average_engagement_time_per_active_user,
    event_count,
    key_events,
    user_key_event_rate,
    total_revenue,
    new_user_ratio,
    low_engagement_rate_flag,
    short_engagement_time_flag,
    review_flag,
    review_reason,
    analysis_start_date,
    analysis_end_date
  )

events_clean <- events_audit |>
  transmute(
    event_name,
    event_count,
    total_users,
    event_count_per_active_user,
    total_revenue,
    event_category,
    tracking_status,
    analysis_start_date,
    analysis_end_date
  )

traffic_clean <- traffic_audit |>
  transmute(
    channel,
    sessions,
    engaged_sessions,
    engagement_rate,
    average_engagement_time_per_session,
    events_per_session,
    event_count,
    key_events,
    session_key_event_rate,
    total_revenue,
    session_share,
    engagement_rate_check,
    engagement_rate_difference,
    channel_flag,
    analysis_start_date,
    analysis_end_date
  )

landing_pages_clean <- landing_page_audit |>
  transmute(
    landing_page_original,
    landing_page_standardised,
    sessions,
    active_users,
    new_users,
    average_engagement_time_per_session,
    key_events,
    total_revenue,
    session_key_event_rate,
    has_query_parameter,
    landing_page_flag,
    analysis_start_date,
    analysis_end_date
  )

pages_clean <- pages_audit |>
  transmute(
    page_path_original,
    page_path_standardised,
    page_path_template,
    page_path_was_templated,
    views,
    active_users,
    views_per_active_user,
    average_engagement_time_per_active_user,
    event_count,
    key_events,
    total_revenue,
    has_query_parameter,
    page_path_flag,
    analysis_start_date,
    analysis_end_date
  )
```

Record every material cleaning or audit rule. `Rows affected` reports the
number of source rows matching the rule; it does not imply that those rows were
deleted.

```{r}
ga4_cleaning_log <- tribble(
  ~step_id, ~dataset, ~field, ~issue_or_rule, ~action,
  ~rows_affected, ~status, ~impact_or_note,

  1, "All GA4 exports", "File metadata",
  "Nine GA4 report metadata rows precede each CSV header",
  "Skipped the first nine rows during import",
  NA_integer_, "Completed",
  "Metadata is excluded from analysis but retained in each raw file",

  2, "All GA4 exports", "Column names",
  "Exported names contain spaces and punctuation",
  "Converted names to snake case with clean_names()",
  NA_integer_, "Completed",
  "Improves reproducibility without changing values",

  3, "Events", "event_name",
  "Event names do not indicate their analytical role",
  "Added event_category and tracking_status",
  nrow(events_clean), "Completed",
  "Original event names and counts are unchanged",

  4, "Traffic Acquisition", "channel",
  "The default-channel-group field name is difficult to use",
  "Renamed the analysis field to channel and added quality flags",
  nrow(traffic_clean), "Completed",
  "No channel was deleted or reassigned",

  5, "Landing Pages", "landing_page_original",
  "Landing-page paths may contain whitespace, query strings or trailing slashes",
  "Created landing_page_standardised for comparison",
  sum(
    landing_pages_clean$landing_page_original !=
      landing_pages_clean$landing_page_standardised,
    na.rm = TRUE
  ),
  "Completed",
  "Original paths remain available",

  6, "Landing Pages", "landing_page_original",
  "GA4 recorded an unidentified landing-page category",
  "Retained (not set) and added an audit flag",
  sum(
    landing_pages_clean$landing_page_original == "(not set)",
    na.rm = TRUE
  ),
  "Retained and flagged",
  "Unidentified sessions were not imputed or removed",

  7, "Pages and Screens", "page_path_original",
  "Paths may contain query strings or trailing slashes",
  "Created page_path_standardised for comparison",
  sum(
    pages_clean$page_path_original !=
      pages_clean$page_path_standardised,
    na.rm = TRUE
  ),
  "Completed",
  "Original paths remain available",

  8, "Pages and Screens", "page_path_template",
  "Dynamic member, order, subscription and group identifiers fragment reporting",
  "Replaced confirmed dynamic IDs and tokens with {id} or {token} in an analysis-only template field",
  sum(pages_clean$page_path_was_templated, na.rm = TRUE),
  "Completed",
  "The website, GA4 configuration and original paths were not changed",

  9, "Pages and Screens", "active_users",
  "Active users are not additive across path variants",
  "Excluded active-user sums from page-template aggregation",
  NA_integer_, "Completed",
  "Prevents double-counting users across multiple paths",

  10, "Demographic", "country",
  "Several high-volume country rows have unusually low engagement",
  "Added transparent review flags; retained all countries",
  sum(demographic_clean$review_flag, na.rm = TRUE),
  "Retained and flagged",
  "A flag is not proof of bot or invalid traffic",

  11, "All GA4 exports", "key_events",
  "All supplied Key Event metrics are zero",
  "Retained the reported zeros and documented the tracking limitation",
  sum(c(
    events_clean$event_count == 0,
    traffic_clean$key_events == 0,
    landing_pages_clean$key_events == 0,
    pages_clean$key_events == 0,
    demographic_clean$key_events == 0
  ), na.rm = TRUE),
  "Not fixable in analysis",
  "Historical conversions cannot be reconstructed reliably from these aggregate exports",

  12, "All GA4 exports", "Source rows",
  "Raw exports must remain the source of truth",
  "No raw row was overwritten or deleted",
  0L, "Completed",
  "All derived data are written to data/clean/GA4"
)

ga4_cleaning_log
```

Write the clean datasets, cleaning log and output manifest to a separate clean
directory. Re-rendering this report reproducibly replaces only these derived
outputs.

```{r}
clean_output_dir <- here("data", "clean", "GA4")

dir.create(
  clean_output_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

clean_outputs <- list(
  ga4_demographic_clean = demographic_clean,
  ga4_events_clean = events_clean,
  ga4_traffic_acquisition_clean = traffic_clean,
  ga4_landing_pages_clean = landing_pages_clean,
  ga4_pages_and_screens_clean = pages_clean
)

iwalk(
  clean_outputs,
  ~ write_csv(
    .x,
    file.path(clean_output_dir, paste0(.y, ".csv")),
    na = ""
  )
)

ga4_output_manifest <- imap_dfr(
  clean_outputs,
  ~ tibble(
    dataset = .y,
    file_name = paste0(.y, ".csv"),
    rows = nrow(.x),
    columns = ncol(.x),
    analysis_start_date = analysis_start_date,
    analysis_end_date = analysis_end_date
  )
)

write_csv(
  ga4_cleaning_log,
  file.path(clean_output_dir, "ga4_cleaning_log.csv"),
  na = ""
)

write_csv(
  ga4_output_manifest,
  file.path(clean_output_dir, "ga4_output_manifest.csv"),
  na = ""
)

ga4_output_manifest
```

## 10. GA4 Audit Summary and Recommendations

### Overall Assessment

The supplied GA4 exports are structurally complete, internally consistent and
suitable for website traffic, content and general engagement analysis. The
main constraints are tracking configuration, attribution detail and path
fragmentation rather than missing CSV values or duplicate source rows.

### Principal Data-Quality Findings

```{r}
ga4_audit_findings <- tibble(
  priority = c(
    "High", "High", "High", "Medium",
    "Medium", "Medium", "Positive control"
  ),
  area = c(
    "Conversion measurement",
    "Campaign attribution",
    "Geographic traffic",
    "Landing-page completeness",
    "Page-path fragmentation",
    "Low-engagement page review",
    "Cross-report reconciliation"
  ),
  finding = c(
    "Form interactions are collected, but all supplied Key Event metrics are zero",
    sprintf(
      "Direct represents %.1f%% of sessions while Email contains only %s sessions",
      100 * tracking_risk_summary$direct_share,
      format(tracking_risk_summary$email_sessions, big.mark = ",")
    ),
    sprintf(
      "%s high-volume countries meet the documented engagement review rules",
      nrow(demographic_review_countries)
    ),
    sprintf(
      "(not set) contains %s sessions (%.1f%% of landing-page sessions)",
      format(landing_page_summary$not_set_sessions, big.mark = ","),
      100 * landing_page_summary$not_set_session_share
    ),
    sprintf(
      "%s dynamic page paths were grouped into %s reporting templates",
      format(page_template_validation$rows_changed_by_templating, big.mark = ","),
      page_template_validation$distinct_templates_created
    ),
    sprintf(
      "%s pages meet the high-view, low-engagement screening rule",
      nrow(high_view_low_engagement_pages)
    ),
    sprintf(
      "Pages and Events reconcile exactly at %s page views and %s total events",
      format(pages_summary$total_views, big.mark = ","),
      format(pages_summary$total_event_count_across_page_rows, big.mark = ",")
    )
  ),
  analytical_impact = c(
    "Channel-, page- and country-level conversion rates cannot be reported reliably",
    "Some campaign traffic may be classified as Direct, limiting channel comparisons",
    "Overall engagement measures are sensitive to the flagged country rows",
    "Entry-page analysis is incomplete for sessions without an identified landing page",
    "Unstandardised URLs would split the same functional page across many rows",
    "The flagged pages require technical or content review before performance conclusions",
    "Confirms that the supplied exports represent a consistent event population"
  )
)

ga4_audit_findings
```

### Cleaning Completed in the Analysis Layer

```{r}
cleaning_completed_summary <- ga4_cleaning_log |>
  filter(status %in% c("Completed", "Retained and flagged")) |>
  select(
    dataset,
    field,
    action,
    rows_affected,
    status,
    impact_or_note
  )

cleaning_completed_summary
```

### Tracking Limitations Not Fixable in This Analysis

```{r}
tracking_limitations <- tribble(
  ~limitation, ~consequence, ~required_follow_up,

  "No Key Events are reported in the supplied exports",
  "Conversion rates cannot be compared across channels, pages or countries",
  "STAV's analytics administrator should review which business outcomes should be configured as Key Events",

  "The Events export does not distinguish individual forms or business outcomes",
  "form_submit cannot be interpreted as event registration or membership conversion",
  "Future tracking should include stable form_name, form_id or business-specific event names",

  "Only the default channel group is available in the acquisition export",
  "Direct, Email and Unassigned traffic cannot be diagnosed in detail",
  "Future exports should include session source, medium, campaign and date",

  "No hostname-by-country or source-by-country breakdown is supplied",
  "Low-engagement overseas traffic cannot be confirmed as automated or irrelevant",
  "Investigate the flagged countries using hostname, source, medium and landing-page dimensions",

  "The supplied exports are aggregate 12-month tables without a date dimension",
  "Monthly trends, campaign timing and before-after comparisons cannot be produced",
  "Include date or month in future recurring exports",

  "User metrics are not additive across countries or page paths",
  "Summed row-level users may double-count people across dimension values",
  "Use GA4 totals or user-level identifiers where an approved, privacy-safe workflow exists",

  "Exit rate and full navigation-path dimensions are not included",
  "High-exit pages and broken journeys cannot be measured from the supplied files",
  "Obtain appropriate GA4 exploration exports for journey analysis"
)

tracking_limitations
```

### Prioritised Recommendations

```{r}
ga4_recommendations <- tribble(
  ~priority, ~recommendation, ~owner, ~expected_benefit,

  "P1", "Define business conversion events and configure approved events as GA4 Key Events",
  "STAV analytics or website administrator",
  "Enables conversion reporting for registrations, membership actions and other outcomes",

  "P1", "Adopt a documented UTM naming convention for email, social and campaign links",
  "STAV marketing team",
  "Improves campaign attribution and reduces avoidable Direct traffic",

  "P1", "Include date, session source, session medium and campaign in future GA4 exports",
  "STAV analytics administrator",
  "Supports monthly reporting and more reliable channel diagnosis",

  "P2", "Investigate (not set) landing pages and reconcile the 1.1% session difference between GA4 reports",
  "STAV analytics or website administrator",
  "Improves landing-page completeness and confidence in report totals",

  "P2", "Review the four flagged countries using hostname, source, medium and landing page before any exclusion decision",
  "Analyst with STAV analytics administrator",
  "Distinguishes relevant audiences from possible automated or unrelated traffic",

  "P2", "Use page templates or a stable content-group/page-type field for member-area reporting",
  "Analyst for reporting; website administrator for future tracking",
  "Prevents dynamic IDs and tokens from fragmenting member engagement analysis",

  "P2", "Technically review high-view pages with near-zero engagement time",
  "STAV website administrator",
  "Identifies redirects, broken pages, tracking gaps or automated traffic",

  "P3", "Use this QMD as the reproducible monthly GA4 audit and cleaning pipeline",
  "Analyst",
  "Creates consistent clean outputs and an auditable processing record"
)

ga4_recommendations
```

### Scope Boundary

All cleaning in this report is performed in the analysis layer. Raw GA4 files
remain unchanged. Changes to WordPress, member-system routes, GA4 event
configuration or campaign-tagging practice are outside the analyst's
implementation scope and are presented only as recommendations for STAV.
